---
title: "Simulation_dataset5"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation_dataset5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MSstatsThermalProfiler)
library(ggplot2)
library(tidyr)
library(MSstats)
library(MSstatsTMT)
```

# Simulations

## Dataset 1-Phaneuf *et al.*

Load in MSstatsTMT processed data

```{r}
#read summarised protein output from MSstatsTMT
summarised.proteins<- readRDS("~/MSstatsThermalProfiler/vignettes/Dataset_5_simulation/summarised.proteinsD1.RDS")
```

Print out ICC distributions and simulate a strong shifter for a design with two temperatures, 2 plexes, 10 replicates

```{r}
#get variance components
Dataset_1_ICC<-msstats_icc(summarised.proteins,temps=unique(summarised.proteins$ProteinLevelData$temperature)[c(5,6)])
#generate simulations
Thermal_2plex_10rep_strong<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
Two_temps_2plex_10rep_strong<-plot_benchmarks_MSstatsTMT(Thermal_2plex_10rep_strong,design="TPP",shifter="strong")
```
Simulate strong interactor with 4 TMT 10-plexes, 4 biological replicates and 10 temperatures

```{r}
#get variance components
Dataset_1_ICC<-msstats_icc(summarised.proteins,temps=unique(summarised.proteins$ProteinLevelData$temperature))
#generate simulations
Thermal_4plex_4strong<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
Ten_temps_4plex_4rep<-plot_benchmarks_MSstatsTMT(Thermal_4plex_4strong,design="TPP",shifter="strong")
```

Simulate weak interactor with 4 TMT 10-plexes, 4 biological replicates and 10 temperatures

```{r}

#generate simulations
Thermal_4plex_4rep_weak_10temps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
Ten_temps_4plex_4rep_weak_10temps<-plot_benchmarks_MSstatsTMT(Thermal_4plex_4rep_weak_10temps,design="TPP",shifter="weak")
```

Strong shifter with two temperatures, 4 plexes, 2 temps, 10 replicates

```{r}
#print out a simulation with two plexes
Thermal_sims_4plex_10reps_strong<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range = seq(5,6),design="TPP")
#save histograms to the workspace along with profile plots
Two_temps_4plex_10reps_strong<-plot_benchmarks_MSstatsTMT_4plex(Thermal_sims_4plex_10reps_strong,design="TPP",shifter="strong")
```
Weak shifter design with two temperatures, 4 plexes, 2 temps, 10 replicates per condition
```{r}
#print out a simulation with two plexes
Thermal_sims_4plex_10reps_weak<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range = seq(5,6),design="TPP")
#save histograms to the workspace along with profile plots
Two_temps_4plex_10reps_weak<-plot_benchmarks_MSstatsTMT_4plex(Thermal_sims_4plex_10reps_weak,design="TPP",shifter="weak")

```

