------------------------------------------------------------------------
title: "Vignette12"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Data_Processing_MSstatsTMT}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
---
#With dataset 3a, this vignette will outline MSstatsTMT processed datasets at the peptide level and evaluate TPP spline, NPARC sigmoid, SCAM splines and MSstatsTMT statistical modeling.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")

```

Library setup

```{r setup}
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(dplyr)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())

```
#With dataset 3a, this vignette will outline Thermal profiling designs at the PSM level and evaluate MSstatsTMT statistical modeling. Thermal profiling designs should use at least two temperatures and more biological replicates would be ideal.  In this case, we focus on a design as shown in Figure 5 panel F with two temperatures in the contrast in a 10-temperature design with 4 biological replicates (2 per condition vehicle and treated).
#A. Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}
WD<-"~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/Reprocessed/"
h<-list.files(WD,pattern="PSMs.txt")|>as.list()
setwd(WD)
df<-purrr::map(h,function(x) read.delim(x)) |> dplyr::bind_rows()

PSMs<-df
df<-as.data.frame(df)
#Filter out shared accessions
#df<-df[!stringr::str_detect(df$`Master Protein Accessions`,";"),]
#Filter out channel 131C
df<-df[,!stringr::str_detect(names(df),"Abundance.131C")]
#rename 131C to 131
names(df)<-stringr::str_replace(names(df),"131N","131")
```

Annotation File Generation

### Option 1: Generate annotation file

```{r}
# 
# 
PD_df<-df|>dplyr::select(Spectrum.File,File.ID,dplyr::starts_with("Abundance"))|>dplyr::mutate(Fraction=stringr::str_remove(File.ID,"F[:digit:]."))|>dplyr::mutate(treatment=ifelse(stringr::str_detect(Spectrum.File,"V|DMSO"),"vehicle","treated"))


PD_df<-PD_df|>tidyr::pivot_longer(cols=colnames(PD_df)[stringr::str_detect(colnames(PD_df),"Abundance")],names_to="Channel",values_to="Abundance")|>dplyr::select(-Abundance)|>dplyr::distinct()|>dplyr::mutate(Channel=stringr::str_extract(Channel,"[[:digit:]]+[N|C]|[:digit:][:digit:][:digit:]"),File.ID=stringr::str_extract(File.ID,"F[:digit:]"),Condition=paste0(Channel,"_",treatment))

Mapping<-PD_df|>dplyr::select(treatment,File.ID)|>dplyr::distinct()|>dplyr::group_by(treatment)|>dplyr::mutate(TechRepMixture=seq(1:dplyr::n()))|>dplyr::ungroup()
# #set mapping between temperatures and channels
temps<-set_temps(10,c("37.3","40.2","44.9","47.6","51","54","57","61.8","64.1","67"))
#TMT11 was used but the 75C temperature was filtered out
# temps$Channel<-ifelse(temps$Channel=="131","131C",temps$Channel)
#add temperatures
df_annotation<-PD_df|>dplyr::select(Spectrum.File,Channel,File.ID,Fraction,Condition)|>dplyr::distinct()|>dplyr::inner_join(temps,by="Channel")|>dplyr::inner_join(Mapping)

#Add bioreplicate
df_annotation$BioReplicate<-ifelse(df_annotation$temperature==min(df_annotation$temperature),"Norm",paste0(df_annotation$Channel,"_",df_annotation$treatment))

df_annotation$Mixture<-paste0(df_annotation$treatment,"_",
                              df_annotation$TechRepMixture)
df_annotation$Run<-df_annotation$Spectrum.File



```



#B. Table 1, steps 1 and 2 
Import the data into MSStatsTMT

```{r}

processed.features <- MSstatsTMT::PDtoMSstatsTMTFormat(df,
                                                       df_annotation,
                                                       which.proteinid = "Master.Protein.Accessions",
                                                       useNumProteinsColumn = FALSE,
                                                       useUniquePeptide = TRUE,
                                                       rmPSM_withfewMea_withinRun = TRUE,
                                                       rmProtein_with1Feature = TRUE,
                                                       summaryforMultipleRows = max,
                                                       use_log_file = FALSE,
                                                       append=FALSE,
                                                       verbose=TRUE)
```
### Table 1 steps 4-6
Summarize the peptide data into proteins

```{r}
processed.features$Condition<-as.character(processed.features$Condition)
processed.features$Condition[stringr::str_detect(processed.features$Condition,"126")]<-"Norm"

summarised.proteins <- MSstatsTMT::proteinSummarization(
  data = processed.features,
  method = "msstats",
  global_norm = FALSE,
  reference_norm = TRUE,
  MBimpute=TRUE,
  use_log_file=FALSE,
  append=FALSE,
  remove_norm_channel=TRUE,
  remove_empty_channel=TRUE)

summarised.proteins$ProteinLevelData<-summarised.proteins$ProteinLevelData|>dplyr::inner_join(temps,by="Channel")
summarised.proteins$FeatureLevelData<-summarised.proteins$FeatureLevelData|>dplyr::inner_join(temps,by="Channel")
summarised.proteins$ProteinLevelData$treatment<-summarised.proteins$ProteinLevelData$Run
summarised.proteins$FeatureLevelData$treatment<-summarised.proteins$FeatureLevelData$Run
```
#C. Table 2

###  Create a matrix to measure the contrast using a subset of temperatures

```{r}
df3<-list()
#copy summarised data to a separate list_df
df3$ProteinLevelData<-summarised.proteins$ProteinLevelData
df3$FeatureLevelData<-summarised.proteins$FeatureLevelData

#choose three temps to contrast around the midpoint

temps_mid<-c("47.6","51","54")

#then create a contrast matrix without the reference channel 

comparison_mid<-make_contrast_matrix(df3,temps=temps_mid)


```

#D.As in Figure 5, panel F to fit linear mixed models with groupComparisons from MSstatsTMT

```{r}


DIM_mid<-groupComparisonThermalProfiling(
  df3,
  contrast.matrix = comparison_mid,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  missing_timepoint = "replace",
  replacement=list("131_vehicle" = c("130C_vehicle"))
)

```


#E. Determine which accessions are kinase targets

```{r}
#read in protein file
WD<-"D:/"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()
setwd(WD)
Proteins<-data.frame()
Proteins<-lapply(h,function(x) readxl::read_xlsx(paste0(WD,x))) |> dplyr::bind_rows()|>dplyr::select(Accession,"Gene Symbol")|>dplyr::rename("GeneSymbol"="Gene Symbol")|>dplyr::distinct()
#Read in human kinome (522 proteins)
kinases<-read.csv(paste0(WD,"/HumanKinome_040920.csv"))
kinases<-kinases$Accession

#join Uniprot IDs to original DF
kinase_df<-kinases|>dplyr::inner_join(Proteins,relationship="many-to-many")|>dplyr::distinct()
```

