---
title: "Vignette9"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calculate_PPV_Sim}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#Figure 4 simulation data is outlined for a template protein in dataset 2 which represents a true negative protein interactor.
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
devtools::load_all(".")
```

```{r setup}
library(scam)
library(mgcv)
library(marginaleffects)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(future)
library(ggplot2)
library(forcats)
library(cowplot)
theme_set(theme_bw())

plan(multicore,workers=future::availableCores()-2)
options(future.globals.maxSize = 8000 * 1024^2)
```

#A.  We begin by simulating true negative values. This simulation contains four replicates, two per condition (e.g. vehicle or treated) and relies on 10 TMT channels where each channel corresponds to one temperature. 

```{r}
data(truenegatives)
truenegatives<-truenegatives$ProteinLevelData|>dplyr::mutate(Accession=Protein,
                                            treatment=Condition,
                                            Mixture=Subject)
```
### Or you can simulate null model that follows a sigmoid function with various instances of biological variance
```{r}
subject_biovar <- c(0.05,0.1,0.2,0.3,0.4)

truenegatives<-purrr::map(subject_biovar, function(x) simulate_multiple_null_sigmoids(n=4, error_sd = 0.05, subject_sd = x)$simdata)|>
  dplyr::bind_rows()

truenegatives<-truenegatives|>
  dplyr::group_by(Group)|>
  dplyr::mutate(Group = as.factor(Group),
                treatment = stringr::str_extract(Group,"[[:lower:]]+"),
                Mixture = Subject)|>
  dplyr::ungroup()

```
#B. TPP splines statistical model
```{r}
start=proc.time()
TPP_F_trueneg_results = TPP_NPARC_calc(truenegatives,method="NPARC",DF=5,temps=set_temps(10,unique(truenegatives$temperature)),
                                        CARRIER=FALSE,returnModels=TRUE,filters=FALSE,NORM = FALSE)
end=proc.time()
print(end-start)
```

###the null distribution of p-values for TPP splines for 5% biological variance

```{r}
TPP5<-ggplot(
  TPP_F_trueneg_results |> filter(stringr::str_detect(Protein, "0.05"))|>dplyr::mutate(BioVar= "% of biovar = 5"),
  aes(x = pNPARC)
) +
  geom_histogram(bins = 40, fill = "#D95F0E",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+ggtitle("TPP stat. model")+
  facet_wrap(~BioVar)+ylim(0,2000)
  
TPP5
```
###the null distribution of p-values for TPP splines for 40% biological variance

```{r}
TPP40<-ggplot(
  TPP_F_trueneg_results |> filter(stringr::str_detect(Protein, "0.4"))|>dplyr::mutate(BioVar= "% of biovar = 40"),
  aes(x = pNPARC)
) +
  geom_histogram(bins = 40, fill = "#D95F0E",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+facet_wrap(~BioVar)+ylim(0,2000)
  

TPP40
```
  
#C. NPARC statistical model

Implement a sigmoidal function for true negatives

```{r}
NPARC_fits_TN <-NPARC::NPARCfit(x = truenegatives$temperature,
                 y =truenegatives$Abundance,
                 id =truenegatives$Protein,
                 groupsNull = NULL,
                 groupsAlt = as.character(truenegatives$Condition),
                 returnModels = TRUE)
```

###NPARC theoretical F-test for 5% biological variance

```{r}
NPARC_test_theo_TN<-NPARC::NPARCtest(NPARC_fits_TN$metrics,dfType="theoretical")

NPARC5<-hist(NPARC_test_theo_TN$pVal[stringr::str_detect(NPARC_test_theo_TN$id,"0.05")],breaks=40,xlab = "p-values",col="#FEC44F",cex.lab = 2,cex.axis=2,main="NPARC stat. model")
```
###NPARC theoretical F-test for 40% biological variance

```{r}
NPARC_test_theo_TN<-NPARC::NPARCtest(NPARC_fits_TN$metrics,dfType="theoretical")

NPARC40<-hist(NPARC_test_theo_TN$pVal[stringr::str_detect(NPARC_test_theo_TN$id,".4")],breaks=40,xlab = "p-values",col="#FEC44F",cex.lab = 2,cex.axis=2,main="")
```

### NPARC empirical F-test for 5% biological variance

```{r}

NPARC_test_emp_TN<-NPARC::NPARCtest(NPARC_fits_TN$metrics,dfType="empirical")

NPARC_5_emp<-ggplot(
  NPARC_test_emp_TN |> filter(stringr::str_detect(id, "0.05"))|>dplyr::mutate(BioVar= "% of biovar = 5"),
  aes(x = pVal)
) +
  geom_histogram(bins = 40, fill = "#FEC44F",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+ggtitle("NPARC stat. model")+facet_wrap(~BioVar)+ylim(0,80)

NPARC_5_emp
```


### NPARC empirical F-test for 40% biological variance

```{r}

NPARC_test_emp_TN<-NPARC::NPARCtest(NPARC_fits_TN$metrics,dfType="empirical")

NPARC40_emp<-ggplot(
  NPARC_test_emp_TN |> filter(stringr::str_detect(id, "0.4"))|>dplyr::mutate(BioVar= "% of biovar = 40"),
  aes(x = pVal)
) +
  geom_histogram(bins = 40, fill = "#FEC44F",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+facet_wrap(~BioVar)+ylim(0,80)

NPARC40_emp

```


#D. SCAM statistical modeling

```{r}
#Add an accession column
truenegatives <- truenegatives |> dplyr::mutate(Accession=Protein)
start=proc.time()
SCAM_pvals_TN = compute_pvalues_ATE_RE_F(truenegatives)
    end=proc.time()
print(end-start)
```

Then look at the null distribution of p-values

```{r}
protein_names<-truenegatives|>dplyr::select(Accession)|>dplyr::distinct()

SCAM_pvals_TN<-cbind(protein_names,SCAM_pvals_TN)

SCAM5<-ggplot(
  SCAM_pvals_TN |> filter(stringr::str_detect(Accession, "0.05"))|>dplyr::mutate(BioVar= "% of biovar = 5"),
  aes(x = p.value)
) +
  geom_histogram(bins = 40, fill = "#00FFFF",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+ggtitle("SCAM stat. model")+facet_wrap(~BioVar)+ylim(0,80)
SCAM5

```

#Then look at the null distribution of p-values

```{r}
SCAM40<-ggplot(
  SCAM_pvals_TN |> filter(stringr::str_detect(Accession, "0.4"))|>dplyr::mutate(BioVar= "% of biovar = 40"),
  aes(x = p.value)
) +
  geom_histogram(bins = 40, fill = "#00FFFF",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+facet_wrap(~BioVar)+ylim(0,80)

SCAM40
```
#E. MSstatsTMT processing
```{r} 
MSstat_TNdata<-list(ProteinLevelData=Sim2MSstatsTMT(truenegatives,half_temp_grid = FALSE))

names(MSstat_TNdata$ProteinLevelData)
```

###Create a contrast matrix for the data

```{r}
MSstat_TNdata$ProteinLevelData<-MSstat_TNdata$ProteinLevelData[!MSstat_TNdata$ProteinLevelData$Channel=="126",]

TN_compmat<-make_contrast_matrix(MSstat_TNdata,temps=unique(MSstat_TNdata$ProteinLevelData$temperature))

TN_compmat
```

#F. MSstatsTMT statistical model

```{r}
start=proc.time()
MSstat_TNresult = MSstatsTMT::groupComparisonTMT(data=MSstat_TNdata,contrast.matrix=TN_compmat,remove_norm_channel = TRUE,
                                     moderated=FALSE,save_fitted_models = FALSE,use_log_file = FALSE)$ComparisonResult


end=proc.time()
print(end-start)

```

#G. For the true negatives from MSstatsTMT, we can calculate the null distribution
### 5% biological variance

```{r}
MSstat_5pctvar<-ggplot(
  MSstat_TNresult |> filter(stringr::str_detect(Protein, "0.05"))|>dplyr::mutate(BioVar= "% of biovar = 5"),
  aes(x = pvalue)
) +
  geom_histogram(bins = 40, fill = "#2C7FB8",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+ggtitle("MSstatsTMT stat. model")+facet_wrap(~BioVar)+ylim(0,80)
MSstat_5pctvar

```
### 40% biological variance
```{r}
MSstat_40pctvar <- ggplot(
  MSstat_TNresult |> filter(stringr::str_detect(Protein, "0.4"))|>dplyr::mutate(BioVar= "% of biovar = 40"),
  aes(x = pvalue)
) +
  geom_histogram(bins = 40, fill = "#2C7FB8",color = 'black')+
  theme(text=element_text(size=16))+xlab("p-value")+facet_wrap(~BioVar)+ylim(0,80)

MSstat_40pctvar
```
```{r}
#generate figure 4

library(ggpubr)
path<-getwd()
histograms_simulation<-ggarrange(plotlist = list(TPP5,NPARC_5_emp,SCAM5,MSstat_5pctvar,TPP40,NPARC40_emp,SCAM40,MSstat_40pctvar),nrow=2,ncol=4)

ggplot2::ggsave(path = path, width = 20, height = 6.5, device='png', dpi=600, filename = "Figure4.png")
```
