---
title: "Data_Processing_MSstatsTMT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
library(devtools)
#devtools::install_version("MASS", version = "7.3-59", repos = "http://cran.us.r-project.org")
# devtools::install_version("Matrix", version = "1.6.2", repos = "http://cran.us.r-project.org")
#devtools::install_github("lme4/lme4",dependencies=TRUE) 

devtools::load_all(".")
```

Library setup

```{r setup}
library(MSstatsTMT)
library(scam)
library(mgcv)
library(marginaleffects)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(R.utils)
library(future)
library(ggplot2)
library(NPARC)
library(TPP)
library(magrittr)
library(ggrepel)
library(readxl)


theme_set(theme_bw())

```

#A. Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}
# WD<-"E:/Zebrafish"
# df<-readPSMs(WD)
```

### Import option 2: Use the example file

**load the peptide spectrum match (PSM) example file into R**

```{r}
#data("PSMsample_HumanData", package = "MSstatsThermalProfiler")

```

### Verify that the PSM data is loaded

```{r}
# df<-PSMsample_HumanData
# head(PSMsample_HumanData)
```

# Import option 3: Read PSM file from folder

```{r}
# Zebrafish_PSMs<-readxl::read_xlsx("E:/Zebrafish/20190508_OR15_AG4_Leijt008_SA_Napa_exp_Control_2_Fraction_Trembl_PSMs.xlsx",.name_repair="universal")

Zebrafish_PSMs<-readxl::read_xlsx("~/work/ivanovlab/figueroa-navedo.a/data/Zebrafish/20190508_OR15_AG4_Leijt008_SA_Napa_exp_Control_2_Fraction_Trembl_PSMs.xlsx",.name_repair= unique)
#rename columns to MSstatsTMT naming 
names(Zebrafish_PSMs)<-stringr::str_replace(names(Zebrafish_PSMs),"# ","X")
names(Zebrafish_PSMs)<-stringr::str_replace(names(Zebrafish_PSMs),"[:punct:](?=)[:space:]","..")
names(Zebrafish_PSMs)<-stringr::str_replace(names(Zebrafish_PSMs),"[:space:]",".")
#data("Zebrafish_PSMs", package = "MSstatsThermalProfiler")
df<-Zebrafish_PSMs
df$Spectrum.File<-stringr::str_remove_all(df$Spectrum.File,"OR15|AG4|Leijt008")
head(df)

```

# Annotation File Generation

### Rename the columns from the PSM file and generate annotation file

**Generate annotation file**

```{r}
pd_annotation<-Annotation2MSstatsTMT(df,solvent="Control",temps=set_temps(10,c(64.0,60.4,57.1,53.8,50.5,47.2,43.9,40.6,37.3,34.0)),reference = "131",CARRIER=FALSE)
```

#B.Table 1, steps 1 and 2 Data processing with MSstatsTMT

Import the data into MSStatsTMT

```{r}


processed.input <- MSstatsTMT::PDtoMSstatsTMTFormat(df,
                                                    pd_annotation,
                                                    which.proteinid = "Master.Protein.Accessions",
                                                    useNumProteinsColumn = FALSE,
                                                    useUniquePeptide = FALSE,
                                                    rmPSM_withfewMea_withinRun = TRUE,
                                                    rmProtein_with1Feature = TRUE,
                                                    summaryforMultipleRows = max,
                                                    use_log_file = FALSE,
                                                    append=FALSE,
                                                    verbose=TRUE)
```
###Table 1, steps 4-6
Summarize the peptide data into proteins

```{r}

                                                   
summarised.proteins <- MSstatsTMT::proteinSummarization(data = processed.input,
                                method = "msstats",
                                global_norm = FALSE,
                                reference_norm = TRUE,
                                MBimpute=TRUE,
                                use_log_file=FALSE,
                                append=FALSE,
                                remove_norm_channel=FALSE,
                                remove_empty_channel=FALSE)


summarised.proteins$ProteinLevelData<-summarised.proteins$ProteinLevelData|>dplyr::inner_join(set_temps(10,c(64.0,60.4,57.1,53.8,50.5,47.2,43.9,40.6,37.3,34.0)))|>dplyr::inner_join(pd_annotation|>dplyr::select(Mixture,treatment)|>dplyr::distinct())
```
#C. Table 2
###  Create a matrix to measure the contrast using a subset of temperatures
    ```{r}
    df3<-list()
    df3$ProteinLevelData<-summarised.proteins$ProteinLevelData |> dplyr::mutate(Channel=as.character(Channel),
                                                                                Mixture=as.character(Mixture),
                                                                                Condition=as.character(Condition))|>as.data.frame()|> dplyr::inner_join(set_temps(10,c(64.0,60.4,57.1,53.8,50.5,47.2,43.9,40.6,37.3,34.0)))|>dplyr::filter(Condition!="Norm")
    
    df3$FeatureLevelData<-summarised.proteins$FeatureLevelData |> dplyr::mutate(Channel=as.character(Channel),
                                                                                Mixture=as.character(Mixture),
                                                                                Condition=as.character(Condition))|>as.data.frame()|> dplyr::inner_join(set_temps(10,c(64.0,60.4,57.1,53.8,50.5,47.2,43.9,40.6,37.3,34.0)))|>dplyr::filter(Condition!="Norm")
       

    #choose three ranges around the midpoint
    temps_late<-c("60.4","57.1","53.8")
    temps_mid<-c("50.5","47.2","43.9")
    temps_ear<-c("47.2","43.9","40.6")
    #then create a contrast matrix without the reference channel 
    comparison_late<-make_contrast_matrix(df3,temps=temps_late)
    comparison_mid<-make_contrast_matrix(df3,temps=temps_mid)
    comparison_ear<-make_contrast_matrix(df3,temps=temps_ear)
      
    ```
###  Fit linear mixed models with groupComparisons from MSstatsTMT
```{r}

DIM_late<-groupComparisonThermalProfiling(
  df3,
  contrast.matrix = comparison_late,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  
)

DIM_mid<-MSstatsTMT::groupComparisonTMT(
  df3,
  contrast.matrix = comparison_mid,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)

DIM_ear<-MSstatsTMT::groupComparisonTMT(
  df3,
  contrast.matrix = comparison_ear,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)
```
### Fit models using filtered data which only considers the variance from the selected contrasts

```{r}
early_data<-list()
middle_data<-list()
later_data<-list()
comparison_early<-comparison_ear[which(!comparison_ear==0)]|>as.matrix()|>t()
colnames(comparison_early)<-colnames(comparison_ear)[which(!comparison_ear==0)]
rownames(comparison_early)<-"DIM"
early_data$ProteinLevelData<-df3$ProteinLevelData[df3$ProteinLevelData$Condition%in%colnames(comparison_early),]
comparison_middle<-comparison_mid[which(!comparison_mid==0)]|>as.matrix()|>t()
rownames(comparison_middle)<-"DIM"
colnames(comparison_middle)<-colnames(comparison_mid)[which(!comparison_mid==0)]
middle_data$ProteinLevelData<-df3$ProteinLevelData[df3$ProteinLevelData$Condition%in%colnames(comparison_middle),]
comparison_later<-comparison_late[which(!comparison_late==0)]|>as.matrix()|>t()
rownames(comparison_later)<-"DIM"
colnames(comparison_later)<-colnames(comparison_late)[which(!comparison_late==0)]
later_data$ProteinLevelData<-df3$ProteinLevelData[df3$ProteinLevelData$Condition%in%colnames(comparison_later),]

DIM_later<-MSstatsTMT::groupComparisonTMT(
later_data,
contrast.matrix = comparison_later,
moderated = FALSE,
adj.method = "BH",
remove_norm_channel = TRUE,
remove_empty_channel = TRUE,
save_fitted_models = TRUE,
use_log_file = FALSE,
append = FALSE,
verbose = TRUE,
log_file_path = NULL
)
DIM_middle<-MSstatsTMT::groupComparisonTMT(
middle_data,
contrast.matrix = comparison_middle,
moderated = FALSE,
adj.method = "BH",
remove_norm_channel = TRUE,
remove_empty_channel = TRUE,
save_fitted_models = TRUE,
use_log_file = FALSE,
append = FALSE,
verbose = TRUE,
log_file_path = NULL
)
DIM_early<-MSstatsTMT::groupComparisonTMT(
early_data,
contrast.matrix = comparison_early,
moderated = FALSE,
adj.method = "BH",
remove_norm_channel = TRUE,
remove_empty_channel = TRUE,
save_fitted_models = TRUE,
use_log_file = FALSE,
append = FALSE,
verbose = TRUE,
log_file_path = NULL
)
```

# Benchmarking

#D. TPP sigmoid

We use TPPtr to perform filtering by R\^2 and plateau

```{r}
#since the MSstats protein summarization output does not have the temperature data, we need to perform an inner join to add this 
proteins<-summarised.proteins$ProteinLevelData |>
  dplyr::mutate(Channel=as.character(Channel),Mixture=as.character(Mixture))|>
  as.data.frame()|>
  dplyr::inner_join(PD_df|>dplyr::select(Channel,temperature)|>dplyr::distinct())



proteins1<-proteins|>dplyr::mutate(Condition=stringr::str_extract(Condition,"[[:lower:]]+"),
shape=ifelse(temperature==min(temperature,na.rm=TRUE),"reference","included"))|>dplyr::distinct()
#Since MSstats outputs are log2 transformed, NPARC runs faster by ratioing the data from all temperatures over the baseline temperature
reps<-pd_annotation|>dplyr::select(Mixture,Subject)|>dplyr::distinct()

proteins<-proteins|>
  as.data.frame()|>
  dplyr::inner_join(reps)
if(length(unique(proteins1$Condition))>2|length(unique(proteins1$Condition)==1)){
  
  proteins1$Condition<-ifelse(stringr::str_detect(stringr::str_extract(stringr::str_remove_all(stringr::str_to_lower(proteins1$Run),"[[:digit:]]+&[[:punct:]]+"),"[[:lower:]]+"),"dmso|control"),"vehicle","treated")
}

proteins1<-proteins1|>
  as.data.frame()|>
  dplyr::group_by(Protein,Run,Mixture)|>
  dplyr::group_split()|>
  lapply(function(x) x|>
  dplyr::mutate(Abundance=2^Abundance))

proteins1<-lapply(proteins1,function(y) y|>as.data.frame()|>dplyr::mutate(Abundance=Abundance/Abundance[1]))|>dplyr::bind_rows()
start=proc.time()
TPPtr_Zebra_results = TPP_NPARC_calc(proteins1,method="TPP",DF=5,CARRIER=FALSE,NORM=FALSE,filters=TRUE,temps=set_temps(10,c(64.0,60.4,57.1,53.8,50.5,47.2,43.9,40.6,37.3,34.0)))
end=proc.time()
print(end-start)
```

#E. TPP splines
```{r}

#Run TPP spline and implement the NPARC F-test with 5 DF
start=proc.time()
TPP_Zebra_results_F = TPP_NPARC_calc(proteins1,method="NPARC",DF=5,CARRIER=FALSE,NORM=FALSE,filters=TRUE,temps=set_temps(10,c(64.0,60.4,57.1,53.8,50.5,47.2,43.9,40.6,37.3,34.0)))
end=proc.time()
print(end-start)
```

#F. NPARC

```{r}

#Then run NPARC
NPARC_Zebra_fits <-NPARC::NPARCfit(x =proteins1$temperature,
                 y =proteins1$Abundance,
                 id =proteins1$Protein,
                 groupsNull = NULL,
                 groupsAlt = as.character(proteins1$Condition),
                 returnModels = TRUE)
```

## Empirical test

```{r}
NPARC_test_emp<-NPARC::NPARCtest(NPARC_Zebra_fits$metrics,dfType="empirical")
hist(NPARC_test_emp$pVal,breaks=20)
```

## Theoretical test

```{r}
NPARC_test_emp<-NPARC::NPARCtest(NPARC_Zebra_fits$metrics,dfType="theoretical")
hist(NPARC_test_emp$pVal,breaks=20)


```

##Rename columns for SCAM

```{r}

Zebra_SCAM_proteins<-MSstatstoSCAM(proteins1)
head(Zebra_SCAM_proteins)

```
#G. SCAM splines
```{r}
DIM_Zebra<-compute_pvalues_ATE_RE_F(Zebra_SCAM_proteins)

```

#H. Figures 1 and 2

###Panel C Row 1

```{r}
Total_prot<-list("TPP"=unique(df$Master.Protein.Accessions),"NPARC"=unique(df$Master.Protein.Accessions),"SCAM"=unique(df$Master.Protein.Accessions),"MSstatsTMT"=unique(df$Master.Protein.Accessions))

Total_prot<-lapply(Total_prot,function(x) as.character(x)|>na.omit())
#Check for shared accessions and collapse isoforms
Total_prot<-lapply(Total_prot,function(x) {
  y<-x[!stringr::str_detect(x,";")]
  y<-unique(stringr::str_remove(y,"-[:digit:]"))
  print(paste0(length(y)," out of ",length(x)," proteins have unique protein groups"))
  return(y)})
#verify if targets are present in the Statistical model fit convergence
Benchmarks_MSstatsTMTproc_targ<-purrr::map(Total_prot,function(x) sum(as.character(x)%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       )))

Total_prot_n<-purrr::map2(Total_prot,Benchmarks_MSstatsTMTproc_targ,function(x,y) data.frame(Proteins=length(unique(as.character(x))),ver_targets=y))|>dplyr::bind_rows()|>dplyr::mutate(Steps="PSMs from PD(unique protein groups):",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```
##Panel C row 2
Then after MSstatsTMT processing

```{r}
Output_proc<-list("TPP"=unique(summarised.proteins$ProteinLevelData$Protein),"NPARC"=unique(summarised.proteins$ProteinLevelData$Protein),"SCAM"=unique(summarised.proteins$ProteinLevelData$Protein),"MSstatsTMT"=unique(summarised.proteins$ProteinLevelData$Protein))


Output_proc<-lapply(Output_proc,function(x) as.character(x)|>na.omit())
Output_proc<-lapply(Output_proc,function(x) {y<-x[!stringr::str_detect(x,";")]
                    y<-unique(stringr::str_remove(y,"-[:digit:]"))
                    return(y)})
#verify if targets are present in the Statistical model fit convergence
Benchmarks_MSstatsTMTproc_targ<-purrr::map(Output_proc,function(x) sum(as.character(x)%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       )))

Output_proc_n<-purrr::map2(Output_proc,Benchmarks_MSstatsTMTproc_targ,function(x,y) data.frame(Proteins=length(unique(as.character(x))),ver_targets=y))|>dplyr::bind_rows()|>dplyr::mutate(Steps="Proteins:\n aggregated after  \n MSstatsTMT proc.(Table 1)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```
###Panel C row 3
Statistical model fit convergence

```{r}
Benchmarks_MSstatsTMTproc<-list("TPP"=unique(TPPtr_Zebra_results$Protein_ID[TPPtr_Zebra_results$model_converged_Treatment_1 & TPPtr_Zebra_results$model_converged_Treatment_2 & TPPtr_Zebra_results$model_converged_Vehicle_1 & TPPtr_Zebra_results$model_converged_Vehicle_2]),"NPARC"=unique(NPARC_test_emp$id),"SCAM"=unique(DIM_Zebra$Accession),"MSstatsTMT"=unique(DIM_late$ComparisonResult$Protein))

Benchmarks_MSstatsTMTproc<-lapply(Benchmarks_MSstatsTMTproc,function(x) as.character(x)|>na.omit())

Benchmarks_MSstatsTMTproc<-lapply(Benchmarks_MSstatsTMTproc,function(x) {y<-x[!stringr::str_detect(x,";")]
                    y<-unique(stringr::str_remove(y,"-[:digit:]"))
                    return(y)})

#verify if targets are present in the Statistical model fit convergence
Benchmarks_MSstatsTMTproc_targ<-purrr::map(Benchmarks_MSstatsTMTproc,function(x) sum(as.character(x)%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       )))

Benchmarks_MSstatsTMTproc_n<-purrr::map2(Benchmarks_MSstatsTMTproc,Benchmarks_MSstatsTMTproc_targ,function(x,y) data.frame(Proteins=length(unique(x)),ver_targets=y))|>dplyr::bind_rows()|>dplyr::mutate(Steps="converged from stat. model\n(Table 2, steps 1-3)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```
###Panel C row 4
For the proteins that passed the R\^2 and plateau filters

```{r}
TPPtr_filters<-unique(TPPtr_Zebra_results$Protein_ID[TPPtr_Zebra_results$passed_filter_Vehicle_1_vs_Treatment_1&TPPtr_Zebra_results$passed_filter_Vehicle_2_vs_Treatment_2])


Benchmarks_MSstatsTMTproc_filt<-purrr::map2(Benchmarks_MSstatsTMTproc,names(Benchmarks_MSstatsTMTproc),function(x,y) ifelse(y=="NPARC",sum(as.character(x) %in% TPPtr_filters),sum(as.character(x)%in%x)))
#verify if targets are present in the filtered data
Benchmarks_MSstatsTMTproc_targ<-
  purrr::map2(Benchmarks_MSstatsTMTproc,names(Benchmarks_MSstatsTMTproc),function(x,y) ifelse(y=="NPARC",sum(as.character(x) %in% TPPtr_filters& as.character(x)%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       )),sum(x%in%x&x%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       ))))

Benchmarks_MSstatsTMTproc_filt_n<-purrr::map2(Benchmarks_MSstatsTMTproc_filt,Benchmarks_MSstatsTMTproc_targ,function(x,y) data.frame(Proteins=x,ver_targets=y))|>dplyr::bind_rows()|>dplyr::mutate(Steps="after post-processing filtering \n(Table 2, steps 1-4)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```
###Panel C row 5
For diferentially expressed proteins

```{r}
Benchmarks_MSstatsTMTproc_p<-list("TPP"=unique(TPPtr_Zebra_results$Protein_ID[TPPtr_Zebra_results$pVal_adj_Vehicle_1_vs_Treatment_1<0.05 & TPPtr_Zebra_results$pVal_adj_Vehicle_2_vs_Treatment_2 < 0.05 & TPPtr_Zebra_results$Protein_ID %in% TPPtr_filters]),"NPARC"=as.character(unique(NPARC_test_emp$id[NPARC_test_emp$pAdj<0.05&NPARC_test_emp$id %in% TPPtr_filters])),"SCAM"=unique(DIM_Zebra$Accession[DIM_Zebra$ATE_padjBH<0.05]),"MSstatsTMT"=unique(DIM_late$ComparisonResult$Protein[DIM_late$ComparisonResult$adj.pvalue<0.05]))

Benchmarks_MSstatsTMTproc_p<-lapply(Benchmarks_MSstatsTMTproc_p,function(x) as.character(x)|>na.omit())

#verify if differentially abundant proteins are present in the filtered data
Benchmarks_MSstatsTMTproc_targ<-
  purrr::map2(Benchmarks_MSstatsTMTproc_p,names(Benchmarks_MSstatsTMTproc_p),function(x,y) ifelse(y=="NPARC",sum(as.character(x) %in% TPPtr_filters&as.character(x)%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       )),sum(x%in%c(#"Q7ZUY3",#histone
                                                                                                                               "Q6NV46",#stat3
                                                                                                                               "Q68SP3",# stat5a,
                                                                          # "A4QNT9",# stat5b,
                                                                           #"Q90Y03",# aldh1a2,
                                                                           "Q0H2G3"# aldh1a3,
                                                                           #"F1QZU7",# aldh1a2.2,
                                                                          # "A2BGR9",# aldh1a2.1,
                                                                           #"F1QLV5",#nqo1,
                                                                           #"F1Q7F3"
                       ))))

Benchmarks_MSstatsTMTproc_p_n<-purrr::map2(Benchmarks_MSstatsTMTproc_p,Benchmarks_MSstatsTMTproc_targ,function(x,y) data.frame(Proteins=length(unique(as.character(x))),ver_targets=y))|>dplyr::bind_rows()|>dplyr::mutate(Steps="differentially abundant \n(Table 2, steps 1-5)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```
##Save R files


```{r}
names(Benchmarks_MSstatsTMTproc_p)<-c("TPP stat. model","NPARC stat. model","SCAM stat. model","MSstatsTMT stat. model")

saveRDS(Benchmarks_MSstatsTMTproc_p,"Zebra_MSstatsTMTprocesed.RDS")
```
#I. Plot the curves Figure 1, panels E-H and save data

```{r}
targets<-summarised.proteins$ProteinLevelData|>as.data.frame()|>dplyr::filter(Protein%in% c("Q6NV46","Q68SP3","Q0H2G3"))|>dplyr::mutate(Accession=Protein)|>dplyr::inner_join(Stat_mod)|>dplyr::group_by(Accession)|>dplyr::group_split()
nam<-as.character(unique(dplyr::bind_rows(targets)$Protein))
target_curves<-lapply(targets,function(x){plotMethods(x,
                                                               temps=c("60.4","57.1","53.8"),
                                                               processing="MSstatsTMT",
                                                               labels=c("E","F","G","H"),fit="Sigmoid")})
names(target_curves)<-nam
#get path
path<-getwd()
target_curves$Q0H2G3
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Curves_MSstatsTMT_sigmoid_Q0H2G3.tiff")

target_curves$Q68SP3
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Curves_MSstatsTMT_sigmoid_Q68SP3.tiff")

target_curves$Q6NV46
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Curves_MSstatsTMT_sigmoid_Q6NV46.tiff")
```

#J. Figure 2, Panel C bar charts

```{r}
bar_Dataset2<-dplyr::bind_rows(Total_prot_n,Output_proc_n,Benchmarks_MSstatsTMTproc_n,Benchmarks_MSstatsTMTproc_filt_n,Benchmarks_MSstatsTMTproc_p_n)



bar_Dataset2$Analysis<-factor(bar_Dataset2$Analysis,levels=c("TPP","NPARC","SCAM","MSstatsTMT"))

col<-data.frame(Analysis=factor(unique(bar_Dataset2$Analysis),levels=c("TPP","NPARC","SCAM","MSstatsTMT")),"stat_model"=c("#D95F0E","#FEC44F","#2C7FB8","#07fff8"))


bar_Dataset2<-bar_Dataset2|>dplyr::inner_join(col)

bar_Dataset2$Analysis<-factor(paste0("MSstatsTMT proc., ",bar_Dataset2$Analysis," stat. model"),levels=c("MSstatsTMT proc., TPP stat. model","MSstatsTMT proc., NPARC stat. model","MSstatsTMT proc., SCAM stat. model","MSstatsTMT proc., MSstatsTMT stat. model"))


bar_Dataset2$Steps<-factor(bar_Dataset2$Steps,levels=c("differentially abundant \n(Table 2, steps 1-5)","after post-processing filtering \n(Table 2, steps 1-4)","converged from stat. model\n(Table 2, steps 1-3)","Proteins:\n aggregated after  \n MSstatsTMT proc.(Table 1)","PSMs from PD(unique protein groups):"))


bar_Dataset2$Protein_targ<-paste0(bar_Dataset2$Proteins," ","(",bar_Dataset2$ver_targets,")")

bDataset2<-ggplot(bar_Dataset2|>dplyr::group_by(Analysis),mapping=aes(x=Steps,y=Proteins,alpha=Steps,fill=stat_model))+geom_col(stat_count=aes(x=Steps,y=Proteins),colour="black")+ylab("Number of proteins remaining after each step")+coord_flip()+facet_wrap(~Analysis,nrow=1, labeller = label_wrap_gen(width=20))+scale_fill_manual(values=c("#2C7FB8","#07fff8","#D95F0E","#FEC44F"))+scale_y_continuous(limits=c(0,18000),breaks = c(0,6000,12000))+theme(text=element_text(size=25,                          color = "black"),axis.text.x = element_text(size=25,face="bold"),axis.text.y=element_text(size=25,face="bold"),strip.text.x=element_text(size=25),legend.position="none", strip.placement = "outside")+xlab("Steps")+geom_text(aes(label=Protein_targ),size=8,position=position_dodge(width=0.2),hjust=-0.05,alpha=0.9)+geom_rect(data = subset(bar_Dataset2, Analysis %in% "MSstatsTMT proc., MSstatsTMT stat. model"), 
                          fill = NA, colour = "green", xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf,size=2)

path = getwd()
bDataset2
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Bar_charts_Tm_MSstatsTMT.tiff")
```

### Venn Diagrams (all methods)

```{r}


Venn_Dataset2data_MSstats<-ggvenn::ggvenn(data=Benchmarks_MSstatsTMTproc_p,fill_color =c("#FEC44F","#D95F0E","#07fff8", "#2C7FB8"),set_name_size=4,stroke_alpha=2,text_size=4,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))
Venn_Dataset2data_MSstats
```

#K. Figure 2, Panel D Venn Diagrams (top 3)

```{r}
png(filename = "Venn_Dataset2data_MSstat3.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_MSstatsTMTproc_p[c(1,2,4)],fill_color =c("#D95F0E","#FEC44F","#2C7FB8"),set_name_size=4,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("MSstatsTMT processed pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```
#L. Figure 2, panel E
##Compare Venn Diagrams between processing

```{r}

#load(Zebra_TPPprocesed,package="MSstatsThermalProfiler")
MSstatsTMT_proc<-lapply(Benchmarks_MSstatsTMTproc_p,function(x) as.character(x)|>na.omit())

TPP_proc<-lapply(Zebra_TPPprocesed,function(x) as.character(x)|>na.omit())


compare_proc<-purrr::map2(TPP_proc,MSstatsTMT_proc,function(x,y) list("MSstatsTMT proc."=y,"TPP proc."=x))

Venn_data<-purrr::map2(compare_proc,names(compare_proc),function(x,y)ggvenn::ggvenn(data=x,fill_color =c("#2C7FB8","white"),set_name_size=0.5,stroke_alpha=2,text_size=,fill_alpha = 0.2,show_percentage = FALSE)+coord_flip())
  
colors1<-as.list(c("#FEC44F","#D95F0E","#07fff8","#2C7FB8"))
  
colors<-list(TPP=c("#D95F0E","#D95F0E"),NPARC=c("#FEC44F","#FEC44F"),SCAM=c("#07fff8","#07fff8"),MSstatsTMT=c("#2C7FB8","#2C7FB8"))

colors<-purrr::set_names(colors,names(colors))
png(filename = "Venn_Dataset2data_processing.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

purrr::map2(compare_proc[1],colors[1],function(x,y)VennDiagram::venn.diagram(
  x = x,
  
  filename = paste0('Venn_Dataset2_processing ',paste("TPP")),
  output = TRUE ,
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 600,
  compression = "lzw",
  lwd = 1,
  cex=0.5,
  cat.cex=0.3,
  rotation.degree=90,
     cat.pos = c(180,0),
          cat.dist = c(0.02, 0.02),
  category.names =names(x),
  col=y,
  scaled=FALSE,
  inverted=FALSE
  
))
  

purrr::map2(compare_proc[2],colors[2],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Dataset2_processing ',paste("NPARC")),
  output = TRUE ,
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 600,
  compression = "lzw",
  lwd = 1,
  col=y,
  cex=0.5,
  cat.cex=0.3,
  rotation.degree=90,
    cat.pos = c(180,0),
    cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))


purrr::map2(compare_proc[3],colors[3],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Dataset2_processing ',paste("SCAM")),
  output = TRUE ,
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 600,
  compression = "lzw",
  lwd = 1,
  col=y,
  cex=0.5,
  cat.cex=0.3,
  rotation.degree=-90,
    cat.pos = c(180,0),
    cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))

purrr::map2(compare_proc[4],colors[4],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Dataset2_processing ',paste("MSstatsTMT")),
  output = TRUE ,
  imagetype="png" ,
  height = 480 , 
  width = 480 , 
  resolution = 600,
  compression = "lzw",
  lwd = 1,
  col=y,
  cex=0.5,
  cat.cex=0.3,
  rotation.degree=-90,
     cat.pos = c(180,0),
          cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))
dev.off()

```
