---
title: "Vignette13"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{Vignette13}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  ```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

```{r setup}
library(MSstatsThermalProfiler)
library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(dplyr)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())
```
##With dataset 3b, this vignette will outline TPP processed datasets at the PSM level.  While the thermal profiling experiment only considered two conditions, the DMSO control and 25x (20μM) Staurosporine, OnePot included additional Staurosporine concentrations (1x, 5x, 10x and 25x Staurosporine, i.e. 5 conditions total), and the number of biological replicates was increased to 3 per condition. Unlike with the conventional thermal profiling, samples from the same biological replicate treated with a total of 10 temperatures within the range of 45-58°C were pooled into one mixture prior to TMT labeling, and allocated to a single TMTpro 16-plex.   This single-plex study did not require a reference channel, and the 16th channel was left empty and discarded from further analyses.

#A. Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}

# h<-list.files(WD,pattern="PSMs")|>as.list()
# 
# df<-purrr::map(h,function(x) readxl::read_xlsx(path=paste0(WD,"/",x),.name_repair = "universal")) |> dplyr::bind_rows()
# head(df)

```

### Import option 2: Read PSM file from folder

```{r}
CETSA_OnePot<-data.table::fread("D:/")
colnames(CETSA_OnePot) <- stringr::str_replace_all(colnames(CETSA_OnePot),"[:punct:]| ",".")

#save original data
df<-OnePot
head(OnePot)
```

#B. Annotation File Import
Read annotation file

```{r}
OnePot_annotation<-data.table::fread("D:/OnePotTPP_annotation.csv")

class(OnePot_annotation$TechRepMixture)<-"numeric"
OnePot_annotation$TechRepMixture<-1
OnePot_annotation$Mixture<-1
OnePot_annotation$Condition<-stringr::str_remove(OnePot_annotation$BioReplicate,"_rep[:digit:]")

```

#C. Data processing with MSstatsTMT

Import the data into MSStatsTMT

```{r}


processed.features <- MSstatsTMT::PDtoMSstatsTMTFormat(OnePot,
                                                       OnePot_annotation,
                                                       which.proteinid = "Master.Protein.Accessions",
                                                       useNumProteinsColumn = FALSE,
                                                       useUniquePeptide = TRUE,
                                                       rmPSM_withfewMea_withinRun = TRUE,
                                                       rmProtein_with1Feature = TRUE,
                                                       summaryforMultipleRows = max,
                                                       use_log_file = FALSE,
                                                       append=FALSE,
                                                       verbose=TRUE)
```
#D. Summarize the peptide data into proteins

```{r}


summarised.proteins <- MSstatsTMT::proteinSummarization(data = processed.features,
                                                        method = "Median",
                                                        global_norm = FALSE,
                                                        reference_norm = FALSE,
                                                        MBimpute=TRUE,
                                                        use_log_file=FALSE,
                                                        append=FALSE,
                                                        remove_norm_channel=FALSE,
                                                        remove_empty_channel=FALSE)
```
#E. Visualize a profile plot for MEK2

```{r}

MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P36507")
```

#F. Fit linear mixed models with groupComparisons from MSstatsTMT

```{r}
DIM_OnePot<-groupComparisonTMT(
  summarised.proteins,
  contrast.matrix = "pairwise",
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)

```
#G. Compare DMSO vs 25x staurosporine as in Figure 7, Panels D-F

```{r}
contrast.matrix<-matrix(nrow=1,ncol=5)
colnames(contrast.matrix)<-unique(df3$ProteinLevelData$Condition)
contrast.matrix[1,]<-c(-1,0,0,0,1)
rownames(contrast.matrix)<-"DMSOvs25x"
DIM_OnePot_DMSOvs25x<-groupComparisonTMT(
  df3,
  contrast.matrix = contrast.matrix,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)
```
#H. Compare DMSO vs all concentrations on avg

```{r}
contrast.matrix<-matrix(nrow=1,ncol=5)
colnames(contrast.matrix)<-unique(df3$ProteinLevelData$Condition)
contrast.matrix[1,]<-c(-1,1/4,1/4,1/4,1/4)
rownames(contrast.matrix)<-"DMSOvsAllAvg"
DIM_OnePot_DMSOvsAllAvg<-groupComparisonThermalProfiling(
  df3,
  contrast.matrix = contrast.matrix,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  missing_timepoint = "replace",
  replacement=list("131_vehicle" = c("130C_vehicle"))
)
```
#I. Read in Human Kinome accessions and remove isoforms
```{r}
WD<-"D:/"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()
setwd(WD)
Proteins<-data.frame()
Proteins<-lapply(h,function(x) readxl::read_xlsx(paste0(WD,x))) |> dplyr::bind_rows()|>dplyr::select(Accession,"Gene Symbol")|>dplyr::rename("GeneSymbol"="Gene Symbol")|>dplyr::distinct()
#Read in human kinome (522 proteins)
kinases<-read.csv(paste0(WD,"/HumanKinome_040920.csv"))

#remove isoforms from UniProt accessions
df<-df|>dplyr::mutate(Accession=stringr::str_remove(Master.Protein.Accessions,"-[[:digit:]]+"))|>dplyr::distinct()


#join Uniprot IDs to original DF
kinase_df<-kinases|>dplyr::inner_join(df,relationship="many-to-many")|>dplyr::select(Accession) |> dplyr::distinct()

```
#J. MSstats interactors are defined 

```{r}
#collect interactions from MSstatsTMT

MSstat_hits<-unique(DIM_OnePot_DMSOvs25x$ComparisonResult$Protein[DIM_OnePot_DMSOvs25x$ComparisonResult$adj.pvalue<0.05])
#remove shared protein accessions
MSstat_hits<-MSstat_hits[!stringr::str_detect(MSstat_hits,";")]

```
