---
title: "Vignette13"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette13}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MSstatsThermalProfiler)
library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(dplyr)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())
```
##With dataset 3b, this vignette will outline TPP processed datasets at the PSM level.  While the thermal profiling experiment only considered two conditions, the DMSO control and 25x (20μM) Staurosporine, OnePot included additional Staurosporine concentrations (1x, 5x, 10x and 25x Staurosporine, i.e. 5 conditions total), and the number of biological replicates was increased to 3 per condition. Unlike with the conventional thermal profiling, samples from the same biological replicate treated with a total of 10 temperatures within the range of 45-58°C were pooled into one mixture prior to TMT labeling, and allocated to a single TMTpro 16-plex.   This single-plex study did not require a reference channel, and the 16th channel was left empty and discarded from further analyses.

#A. Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}
WD<-"D:/"
h<-list.files(WD,pattern="PSMs.txt")|>as.list()
setwd(WD)
df<-purrr::map(h,function(x) read.delim(x)) |> dplyr::bind_rows()

PSMs<-df
df<-as.data.frame(df)
#Filter out channel 131C
df<-df[,!stringr::str_detect(names(df),"Abundance.131C")]
#rename 131C to 131
names(df)<-stringr::str_replace(names(df),"131N","131")
```

Annotation File Generation

### Option 1: Generate annotation file

```{r}
 
PD_df<-df|>dplyr::select(Spectrum.File,File.ID,dplyr::starts_with("Abundance"))|>dplyr::mutate(Fraction=stringr::str_remove(File.ID,"F[:digit:]."))|>dplyr::mutate(treatment=ifelse(stringr::str_detect(Spectrum.File,"V|DMSO"),"vehicle","treated"))


PD_df<-PD_df|>tidyr::pivot_longer(cols=colnames(PD_df)[stringr::str_detect(colnames(PD_df),"Abundance")],names_to="Channel",values_to="Abundance")|>dplyr::select(-Abundance)|>dplyr::distinct()|>dplyr::mutate(Channel=stringr::str_extract(Channel,"[[:digit:]]+[N|C]|[:digit:][:digit:][:digit:]"),File.ID=stringr::str_extract(File.ID,"F[:digit:]"),Condition=paste0(Channel,"_",treatment))

Mapping<-PD_df|>dplyr::select(treatment,File.ID)|>dplyr::distinct()|>dplyr::group_by(treatment)|>dplyr::mutate(TechRepMixture=seq(1:dplyr::n()))|>dplyr::ungroup()
# #set mapping between temperatures and channels
temps<-set_temps(10,c("37.3","40.2","44.9","47.6","51","54","57","61.8","64.1","67"))
#TMT11 was used but the 75C temperature was filtered out

#add temperatures
df_annotation<-PD_df|>dplyr::select(Spectrum.File,Channel,File.ID,Fraction,Condition)|>dplyr::distinct()|>dplyr::inner_join(temps,by="Channel")|>dplyr::inner_join(Mapping)

#Add bioreplicate, mixture and run
df_annotation$BioReplicate<-ifelse(df_annotation$temperature==min(df_annotation$temperature),"Norm",paste0(df_annotation$Channel,"_",df_annotation$treatment))

df_annotation$Mixture<-paste0(df_annotation$treatment,"_",
                                           df_annotation$TechRepMixture)
df_annotation$Run<-df_annotation$Spectrum.File

```



#B. Table 1, steps 1 and 2 
Import the data into MSStatsTMT

```{r}

processed.features <- MSstatsTMT::PDtoMSstatsTMTFormat(df,
                                                    df_annotation,
                                                    which.proteinid = "Master.Protein.Accessions",
                                                    useNumProteinsColumn = FALSE,
                                                    useUniquePeptide = TRUE,
                                                    rmPSM_withfewMea_withinRun = TRUE,
                                                    rmProtein_with1Feature = TRUE,
                                                    summaryforMultipleRows = max,
                                                    use_log_file = FALSE,
                                                    append=FALSE,
                                                    verbose=TRUE)
```
### Table 1 steps 4-6
Summarize the peptide data into proteins

```{r}
processed.features$Condition<-as.character(processed.features$Condition)
processed.features$Condition[stringr::str_detect(processed.features$Condition,"126")]<-"Norm"

summarised.proteins <- MSstatsTMT::proteinSummarization(
  data = processed.features,
  method = "msstats",
  global_norm = FALSE,
  reference_norm = TRUE,
  MBimpute=TRUE,
  use_log_file=FALSE,
  append=FALSE,
  remove_norm_channel=TRUE,
  remove_empty_channel=TRUE)

summarised.proteins$ProteinLevelData<-summarised.proteins$ProteinLevelData|>dplyr::inner_join(temps,by="Channel")
summarised.proteins$FeatureLevelData<-summarised.proteins$FeatureLevelData|>dplyr::inner_join(temps,by="Channel")
summarised.proteins$ProteinLevelData$treatment<-summarised.proteins$ProteinLevelData$Run
summarised.proteins$FeatureLevelData$treatment<-summarised.proteins$FeatureLevelData$Run
```
#C. Table 2

###  Create a matrix to measure the contrast using a subset of temperatures

```{r}
df3<-list()
#copy summarised data to a separate list_df
df3$ProteinLevelData<-summarised.proteins$ProteinLevelData
df3$FeatureLevelData<-summarised.proteins$FeatureLevelData

#choose three temps to contrast around the midpoint

temps_mid<-c("47.6","51","54")

#then create a contrast matrix without the reference channel 

comparison_mid<-make_contrast_matrix(df3,temps=temps_mid)


```

#D. As in Figure 5, panel E Fit models using filtered data which only considers the variance from the selected contrasts
```{r}

middle_data<-list()
comparison_middle<-comparison_mid[which(!comparison_mid==0)]|>as.matrix()|>t()
rownames(comparison_middle)<-"DIM"
colnames(comparison_middle)<-colnames(comparison_mid)[which(!comparison_mid==0)]

middle_data$ProteinLevelData<-df3$ProteinLevelData[df3$ProteinLevelData$Condition%in%colnames(comparison_middle),]


```

#E. As in Figure 5, panel F to fit linear mixed models with groupComparisons from MSstatsTMT

```{r}

DIM_middle<-MSstatsTMT::groupComparisonTMT(
middle_data,
contrast.matrix = comparison_middle,
moderated = FALSE,
adj.method = "BH",
remove_norm_channel = TRUE,
remove_empty_channel = TRUE,
save_fitted_models = TRUE,
use_log_file = FALSE,
append = FALSE,
verbose = TRUE,
log_file_path = NULL
)

```
