------------------------------------------------------------------------

---
title: "Data_Processing_MSstatsTMT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")

```

Library setup

```{r setup}

library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())

```

# Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}
CETSA_OnePot<-data.table::fread("D:/PhD Dissertation/Dataset_3b_OnePot/OnePotTPPStauro_Fx_PSMs.txt")
colnames(CETSA_OnePot) <- stringr::str_replace_all(colnames(CETSA_OnePot),"[:punct:]| ",".")

#save original data
df<-CETSA_OnePot
head(CETSA_OnePot)

```


# Annotation File Generation

## Read annotation file

```{r}
CETSA_OnePot_annotation<-data.table::fread("D:/PhD Dissertation/Dataset_3b_OnePot/OnePotTPP_annotation.csv")

class(CETSA_OnePot_annotation$TechRepMixture)<-"numeric"
CETSA_OnePot_annotation$TechRepMixture<-1
CETSA_OnePot_annotation$Mixture<-1

CETSA_OnePot_annotation$Condition<-stringr::str_remove(CETSA_OnePot_annotation$BioReplicate,"_rep[:digit:]")

```

# Data processing with MSstatsTMT

1.  Import the data into MSStatsTMT

```{r}


processed.features <- MSstatsTMT::PDtoMSstatsTMTFormat(CETSA_OnePot,
                                                    CETSA_OnePot_annotation,
                                                    which.proteinid = "Master.Protein.Accessions",
                                                    useNumProteinsColumn = FALSE,
                                                    useUniquePeptide = TRUE,
                                                    rmPSM_withfewMea_withinRun = TRUE,
                                                    rmProtein_with1Feature = TRUE,
                                                    summaryforMultipleRows = max,
                                                    use_log_file = FALSE,
                                                    append=FALSE,
                                                    verbose=TRUE)
```

2.  Summarize the peptide data into proteins

```{r}

                                                   
summarised.proteins <- MSstatsTMT::proteinSummarization(data = processed.features,
                                method = "Median",
                                global_norm = TRUE,
                                reference_norm = FALSE,
                                MBimpute=TRUE,
                                use_log_file=FALSE,
                                append=FALSE,
                                remove_norm_channel=FALSE,
                                remove_empty_channel=FALSE)

#remove shared accessions and collapse isoforms
summarised.proteins<-lapply(summarised.proteins,function(x) {
  if(any(names(x)=="Protein")){
    x$Protein<-as.character(x$Protein)
    x$Protein<-stringr::str_remove_all(x$Protein,"-[:digit:]")
    x<-x[!stringr::str_detect(x$Protein,";"),]
  }else if(any(names(x)=="ProteinName")){
    x$ProteinName<-as.character(x$ProteinName)
    x$ProteinName<-stringr::str_remove_all(x$ProteinName,"-[:digit:]")
    x<-x[!stringr::str_detect(x$ProteinName,";"),]
  }
  return(x)
})
```

3.  Visualize a profile plot for MEK2

    ```{r}

    MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P36507")
    ```

4.  Create a contrast matrix to measure the DIM

    ```{r}
    df3<-list()

    df3$ProteinLevelData<-summarised.proteins$ProteinLevelData

    ```

5.  Fit linear mixed models with groupComparisons from MSstatsTMT

```{r}
DIM_OnePot<-groupComparisonThermalProfiling(
  summarised.proteins,
  contrast.matrix = "pairwise",
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  missing_timepoint = "replace",
  replacement=list("131_vehicle" = c("130C_vehicle"))
)

```

# Compare DMSO vs 25x staurosporine

```{r}
contrast.matrix<-matrix(nrow=1,ncol=5)
colnames(contrast.matrix)<-unique(df3$ProteinLevelData$Condition)
contrast.matrix[1,]<-c(-1,0,0,0,1)
rownames(contrast.matrix)<-"DMSOvs25x"
DIM_OnePot_DMSOvs25x<-groupComparisonThermalProfiling(
  df3,
  contrast.matrix = contrast.matrix,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  missing_timepoint = "replace",
  replacement=list("131_vehicle" = c("130C_vehicle"))
)
```

# Compare DMSO vs all concentrations on average

```{r}
contrast.matrix<-matrix(nrow=1,ncol=5)
colnames(contrast.matrix)<-unique(df3$ProteinLevelData$Condition)
contrast.matrix[1,]<-c(-1,1/4,1/4,1/4,1/4)
rownames(contrast.matrix)<-"DMSOvsAllAvg"
DIM_OnePot_DMSOvsAllAvg<-groupComparisonThermalProfiling(
  df3,
  contrast.matrix = contrast.matrix,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  missing_timepoint = "replace",
  replacement=list("131_vehicle" = c("130C_vehicle"))
)
```

## Get Kinase compiled by Xu *et. al.*

```{r}
WD<-"~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()
setwd(WD)
Proteins<-data.frame()
Proteins<-lapply(h,function(x) readxl::read_xlsx(paste0(WD,"/",x))) |> dplyr::bind_rows()|>dplyr::rename(GeneSymbol="Gene Symbol")|>dplyr::select(Accession,GeneSymbol)|>dplyr::distinct()

kinases<-readxl::read_xlsx(paste0(WD,"/Kinase accessions.xlsx"))|>dplyr::inner_join(Proteins,relationship="many-to-many")

kinase_df<-df|>dplyr::mutate(Accession=Master.Protein.Accessions)|>dplyr::inner_join(kinases,by="Accession",relationship="many-to-many")
#join geneSymbols to original DF
kinase_df<-kinase_df|>dplyr::inner_join(Proteins,relationship="many-to-many")
kinases<-kinase_df$Accession[kinase_df$GeneSymbol %in% kinase_df$GeneSymbol]
kinases<-unique(kinases)

```

## Prepare Venn Diagrams for kinase ID benchmarks

```{r}

#Published work
OnePotTPP_2D_hits_kinases<-unique(kinase_df$Accession[stringr::str_detect(kinase_df$`DetectedOnePotTPP?`,"Y")&kinase_df$Accession%in% kinases])|>na.omit()

SPROX_2D_hits_kinases<-unique(kinase_df$Accession[stringr::str_detect(kinase_df$`DetectedOnePotSPROX?`,"Y")&kinase_df$Accession%in% kinases])|>na.omit()

OnePot_SPROX_hits_kinases<-unique(kinase_df$Accession[stringr::str_detect(kinase_df$`OnePotSPROXHits?`,"Y")&kinase_df$Accession%in% kinases])|>na.omit()

TPP_hits_kinases<-unique(kinase_df$Accession[stringr::str_detect(kinase_df$`OnePotTPPHits?`,"Y")&kinase_df$Accession%in% kinases])|>na.omit()

MSstat_hits<-unique(DIM_OnePot_DMSOvs25x$ComparisonResult$Protein[DIM_OnePot_DMSOvs25x$ComparisonResult$adj.pvalue<0.05])

Benchmarks_external<-list("OnePot 2D TPP"=OnePotTPP_2D_hits_kinases,
                          "MSstatsTMT"=MSstat_hits)

png(filename = "Venn_TPPpapervsThiswork_DMSOvs25x_MSstatsTMTproc_Median.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#cb9c3f","#FEC44F"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```


# Compare full curve TPP vs OnePot and OnePot MSstatsTMT

NPARC

```{r}
names(Thermal_MSstatsTMTprocesed)<-c("NPARC temp.","TPP temp.","SCAM temp.","MSstatsTMT temp.")

MSstat_hits<-unique(stringr::str_remove(as.character(DIM_OnePot_DMSOvs25x$ComparisonResult$Protein[DIM_OnePot_DMSOvs25x$ComparisonResult$adj.pvalue<0.05]),"-[:digit:]"))
MSstat_hits<-MSstat_hits[!stringr::str_detect(MSstat_hits,";")]
#add benchmarks to original full curve designs
Benchmarks_external<-append(list("NPARC temp."=Thermal_MSstatsTMTprocesed$`NPARC temp.`),list(
  "kinases"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennNPARC_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#FEC44F","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```

TPP

```{r}
Benchmarks_external<-append(list("TPP temp."=Thermal_MSstatsTMTprocesed$`TPP temp.`),list(
  "kinases"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennTPP_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#D95F0E","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```

SCAM

```{r}
Benchmarks_external<-append(list("SCAM temp."=Thermal_MSstatsTMTprocesed$`SCAM temp.`),list(
  "kinases"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennSCAM_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#07fff8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```

MSstatsTMT

```{r}
Benchmarks_external<-append(list("MSstatsTMT temp."=Thermal_MSstatsTMTprocesed$`MSstatsTMT temp.`),list(
  "kinases"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennMSstatsTMT_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#2c7fb8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```

# Plot profile plots

```{r}
#Rename columns for Conditions

summarised.proteins<-lapply(summarised.proteins,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","A_DMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","B_1xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","C_5xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","D_10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","E_25xstauro",x$Condition)
                            return(x)})
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P36507")
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein=  "Q9Y5S2")
```

# Venn Diagrams Protein vs Peptide-level

```{r}
OnePot_MSstatsTMTproc_unique_hits <- readRDS("~/work/ivanovlab/figueroa-navedo.a/MSstatsThermalProfiler/OnePot_MSstatsTMTproc_unique_hits.RDS")
OnePot_TPPproc_hits <- readRDS("~/work/ivanovlab/figueroa-navedo.a/MSstatsThermalProfiler/OnePot_TPPproc_hits.RDS")

Benchmarks_external<-list("MSstatsTMT proc."=OnePot_MSstatsTMTproc_unique_hits,
                                 "STRING targets"=na.omit(kinases),
                          "Protein-level proc." = OnePot_TPPproc_hits
                                 )

png(filename = "VennProtein_vs_Peptide_OnePot.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#1d547a","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```
