---
title: "Data_Processing_MSstatsTMT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")
```

Library setup

```{r setup}

library(scam)
library(mgcv)
library(marginaleffects)
library(dplyr)
library(tidyr)
library(purrr)
library(furrr)
library(R.utils)
library(future)
library(ggplot2)
library(NPARC)
library(TPP)
library(magrittr)
library(ggrepel)

theme_set(theme_bw())

```

# Data Import

### Import option 1: Read in your own file

```{r}
WD<-"~/work/ivanovlab/figueroa-navedo.a/data/Human"
#WD<-"E:/internal_dataset/PD_files"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()

PD_df<-MSstatsThermalProfiler::read_cetsa(WD,WD,Frac=TRUE,temps=set_temps(11,c(37.3, 40.6, 43.9, 47.2, 50.5, 53.8, 57.1, 60.4, 64, 67,68)),solvent="DMSO",CARRIER=TRUE)
head(PD_df)
```
###Generate annotation file
```{r}

PD_df$Condition<-as.character(PD_df$Condition)
Mapping<-PD_df|>dplyr::ungroup()|>dplyr::select(File.ID,Spectrum.File,Condition)|>dplyr::distinct()|>
  dplyr::group_by(Condition)|>
  dplyr::arrange(Condition)|>
  dplyr::mutate(
    treatment=ifelse(stringr::str_detect(Spectrum.File,"DMSO"),"vehicle","treated"),
                TechRepMixture=seq(1,dplyr::n()))|>
  dplyr::ungroup()

# #set mapping between temperatures and channels
temps<-set_temps(11,c(37.3, 40.6, 43.9, 47.2, 50.5, 53.8, 57.1, 60.4, 64, 67,68))
#TMT11 was used but the 75C temperature was filtered out
# temps$Channel<-ifelse(temps$Channel=="131","131C",temps$Channel)
#add temperatures
df_annotation<-PD_df|>dplyr::select(Spectrum.File,Channel,File.ID)|>dplyr::distinct()|>dplyr::inner_join(temps,by="Channel")|>dplyr::inner_join(Mapping)

#Add bioreplicate
df_annotation$BioReplicate<-ifelse(df_annotation$temperature==min(df_annotation$temperature),"Norm",paste0(df_annotation$Channel,"_",df_annotation$treatment))

df_annotation$Mixture<-paste0(df_annotation$treatment,"_",
                                           df_annotation$TechRepMixture)





df_annotation$Run<-df_annotation$Spectrum.File

#Filter out shared accessions
#df<-df[!stringr::str_detect(df$`Master Protein Accessions`,";"),]
#Filter out channel 131N
PD_df<-PD_df[!stringr::str_detect(PD_df$Channel,"131C"),]
#replace 131C with 131
PD_df$Channel<-ifelse(PD_df$Channel=="131N","131",PD_df$Channel)

PD_df$treatment<-stringr::str_extract(PD_df$Condition,"[[:lower:]]+")


#same for annotation
df_annotation<-df_annotation[!stringr::str_detect(df_annotation$Channel,"131C"),]
#replace 131C with 131
df_annotation$Channel<-ifelse(df_annotation$Channel=="131N","131",df_annotation$Channel)

#set bioreplicate
df_annotation$BioReplicate<-paste0(df_annotation$treatment,"_",df_annotation$TechRepMixture)



```
###Get a sense of missing value distributions

```{r}

mapping <-PD_df |>dplyr::select(Condition,Subject)|>
  dplyr::mutate(Condition=as.character(Condition))|>
  dplyr::distinct()|>
  dplyr::group_by(Condition)|>
  dplyr::arrange(Condition)|>
  dplyr::mutate(TechRepMixture=seq(1,dplyr::n()),
                Color=ifelse(Condition=="vehicle"&TechRepMixture==1,"#00BCF4",
                             ifelse(Condition == "vehicle" & TechRepMixture==2,"#2b8cbe",
                                    ifelse(Condition=="treated" & TechRepMixture==1,"#F8776D",
                                            "#fa9fb5"))))|>
  dplyr::ungroup()

MV<-PD_df|>
  dplyr::inner_join(mapping)|>
  dplyr::filter(TechRepMixture==1|TechRepMixture==2)|>
  dplyr::group_by(Accession,Subject)|>
  dplyr::select(Accession,Subject,Channel,temperature,Abundance,Condition,TechRepMixture,Color)|>
  dplyr::distinct()|>
  dplyr::filter(Channel!="131C")|>
  dplyr::group_split()

MV<-lapply(MV,function(x) x|>
  dplyr::mutate(missing_channels=sum(is.na(Abundance)),
                plex = paste0(Condition,"_",TechRepMixture))|>
  dplyr::select(Accession,missing_channels,plex,Condition,Color,TechRepMixture)|>
  dplyr::distinct())|>dplyr::bind_rows()
  

histogram_mv<-ggplot(MV,mapping=aes(x=missing_channels,fill=plex),group=Condition)+geom_histogram()+facet_wrap(~Condition)+xlab("# of missing channels")+xlim(0,10)+
  scale_x_continuous(breaks=seq(0,10,1),labels=seq(0,10,1),limits= c(-1,11))+
  ylab("# of Proteins")+
  scale_fill_manual(values = mapping$Color)+theme(text=element_text(size =20))
  
  
path = getwd()
histogram_mv
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "MV_Dataset1.tiff")

```

###Only for dataset 1: filter by missing values

```{r}
PD_df_filtered<-PD_df |>
  dplyr::inner_join(MV|>
                      dplyr::select(Accession,plex,missing_channels,TechRepMixture)|>
                      dplyr::rename(Experiment=plex)|>
                      dplyr::distinct())|>
  dplyr::filter(missing_channels<=3)
proteins_missing<-length(unique(PD_df$Accession))-length(unique(PD_df_filtered$Accession))
print(paste0(round(proteins_missing/length(unique(PD_df$Accession))*100,2),"% of proteins were filtered out with more than 3 missing values"))

```

###Annotation file

```{r}

# #set mapping between temperatures and channels
temps<-set_temps(11,c(37.3, 40.6, 43.9, 47.2, 50.5, 53.8, 57.1, 60.4, 64, 67,68))
#Add bioreplicate
df_annotation$BioReplicate<-ifelse(df_annotation$temperature==min(df_annotation$temperature),"Norm",paste0(df_annotation$Channel,"_",df_annotation$treatment))

df_annotation$Mixture<-paste0(df_annotation$treatment,"_",
                                           df_annotation$TechRepMixture)


df_annotation$Condition<-ifelse(df_annotation$temperature==min(df_annotation$temperature),"Norm",paste0(df_annotation$Channel,"_",df_annotation$treatment))


df_annotation$Run<-df_annotation$Spectrum.File

#Filter out shared accessions
#df<-df[!stringr::str_detect(df$`Master Protein Accessions`,";"),]
PD_df<-PD_df|>
  dplyr::inner_join(MV|>dplyr::select(Accession,plex,missing_channels)|>
                      dplyr::rename(Experiment=plex)|>
                      dplyr::distinct())|>
  dplyr::filter(missing_channels <=3)

#replace 131C with 131
PD_df$Channel<-ifelse(PD_df$Channel=="131N","131",PD_df$Channel)

#same for annotation
df_annotation<-df_annotation[!stringr::str_detect(df_annotation$Channel,"131C"),]
#replace 131C with 131
df_annotation$Channel<-ifelse(df_annotation$Channel=="131N","131",df_annotation$Channel)
df_annotation$Condition<-df_annotation$treatment

PD_df<-PD_df|>dplyr::inner_join(df_annotation) |> dplyr::mutate(Protein = Accession)
```

#### Normalize the data

```{r}
Human_TPPnorm_Proteins<-TPP_normalization(PD_df_filterede,TPPfilters=FALSE,temps=set_temps(10,c(37.3, 40.6, 43.9, 47.2, 50.5, 53.8, 57.1, 60.4, 64, 67)),reference="126",CARRIER=FALSE)

    Human_TPPnorm_Proteins$normData<-Human_TPPnorm_Proteins$normData
proteins<-list(ProteinLevelData=Human_TPPnorm_Proteins$normData|>as.data.frame()|>dplyr::mutate(treatment=stringr::str_extract(Experiment,"[[:lower:]]+")))


```

## Add comparison

```{r}
df<-list(ProteinLevelData=PD_df_filtered,FeatureLevelData=NA)

comparison<-make_contrast_matrix(df,temps=c("53.8","57.1","60.4"))
comparison_mid<-make_contrast_matrix(df,temps=c( "47.2", "50.5", "53.8"))
comparison
```

6.  Fit linear mixed models with groupComparisons from MSstatsTMT

```{r}
DIM<-MSstatsThermalProfiler::groupComparisonThermalProfiling(
  df,
  contrast.matrix = comparison,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)

DIM_mid<-MSstatsThermalProfiler::groupComparisonThermalProfiling(
  df,
  contrast.matrix = comparison_mid,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)
```

# Benchmarking

## TPPtr

benchmarking sigmoid with Tm stat model

```{r}
proteins1<-proteins$ProteinLevelData|>dplyr::mutate(treatment=stringr::str_remove(stringr::str_extract(stringr::str_to_lower(Run),"[[:lower:]]+_"),"_"),
                                              Protein=as.character(Protein),
                                              replicate=stringr::str_extract(TechRepMixture,"[[:digit:]]+"),
                                              treatment=as.character(treatment))
proteins1<-proteins1|>data.frame()
#Run TPP spline and implement the NPARC F-test with 5 DF
start=proc.time()
TPPtr_human_results = TPP_NPARC_calc(proteins1,method="TPP",DF=5,CARRIER=TRUE,NORM=FALSE,filters=TRUE,temps=set_temps(10,c(37.3, 40.6, 43.9, 47.2, 50.5, 53.8, 57.1, 60.4, 64, 67))) 
end=proc.time()
print(end-start)
```

## TPP

benchmarking splines with Ftest

```{r}

#Run TPP spline and implement the NPARC F-test with 5 DF
start=proc.time()
TPP_Human_results_F = TPP_NPARC_calc(proteins1,method="NPARC",DF=5,CARRIER=TRUE,NORM=FALSE,temps=set_temps(10,c(37.3, 40.6, 43.9, 47.2, 50.5, 53.8, 57.1, 60.4, 64, 67)))
end=proc.time()
print(end-start)
```

## NPARC

benchmarking sigmoids with Ftest

```{r}

#Then run NPARC
NPARC_human_fits <-NPARC::NPARCfit(x=proteins1$temperature,
                               y =proteins1$Abundance,
                               id =proteins1$Protein,
                               groupsNull = NULL,
                               groupsAlt = as.character(proteins1$treatment),
                               returnModels = TRUE)

```

## Empirical test

```{r}
NPARC_test_emp<-NPARC::NPARCtest(NPARC_human_fits$metrics,dfType="empirical")
hist(NPARC_test_emp$pAdj,breaks=20)
```

## Theoretical test

```{r}
NPARC_test_theo<-NPARC::NPARCtest(NPARC_human_fits$metrics,dfType="theoretical")
hist(NPARC_test_theo$pAdj,breaks=20)


```

## SCAM Splines

```{r}

Human_SCAM_proteins<-MSstatstoSCAM(proteins1)
head(Human_SCAM_proteins)

```

```{r}
DIM_Human_SCAM<-compute_pvalues_ATE_RE_F(Human_SCAM_proteins)
  
```

# Bar charts

For the total number of proteins

```{r}
Total_prot<-list("TPP"=unique(Human_TPPnorm_Proteins$normData$Protein),"NPARC"=unique(Human_TPPnorm_Proteins$normData$Protein),"SCAM"=unique(Human_TPPnorm_Proteins$normData$Protein),"MSstatsTMT"=unique(Human_TPPnorm_Proteins$normData$Protein))
#remove shared accessions
Total_prot<-lapply(Total_prot,function(x) x[!stringr::str_detect(x,";")])
#remove missing values
Total_prot<-lapply(Total_prot,function(x) as.character(x)|>na.omit())

#verify if targets are present in the Statistical model fit convergence
Benchmarks_TPPproc_targ<-purrr::map(Total_prot,function(x) sum(as.character(x)%in%c("P36507","Q02750")))

Total_prot_n<-purrr::map2(Total_prot,Benchmarks_TPPproc_targ,function(x,y) data.frame(Proteins=as.character(length(unique(x))),ver_targets=as.character(y)))|>dplyr::bind_rows()|>dplyr::mutate(Steps="Proteins (unique protein groups):\n from PD",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```

output from TPP processing

```{r}
Output_proc<-list("TPP"=unique(Human_TPPnorm_Proteins$normData$Protein),"NPARC"=unique(Human_TPPnorm_Proteins$normData$Protein),"SCAM"=unique(Human_TPPnorm_Proteins$normData$Protein),"MSstatsTMT"=unique(Human_TPPnorm_Proteins$normData$Protein))
#remove shared accessions
Output_proc<-lapply(Output_proc,function(x) x[!stringr::str_detect(x,";")])
#remove missing values
Output_proc<-lapply(Output_proc,function(x) as.character(x)|>na.omit())

#verify if targets are present in the Statistical model fit convergence
Benchmarks_TPPproc_targ<-purrr::map(Output_proc,function(x) sum(x%in%c("P36507","Q02750")))

Output_proc_n<-purrr::map2(Output_proc,Benchmarks_TPPproc_targ,function(x,y) data.frame(Proteins=as.character(length(unique(x))),ver_targets=as.character(y)))|>dplyr::bind_rows()|>dplyr::mutate(Steps="Proteins: after \n TPP normalization (Table 1)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```

Statistical model fit convergence

```{r}
Benchmarks_TPPproc<-list(
  "TPP"=unique(TPPtr_human_results$Protein_ID[TPPtr_human_results$model_converged_treated_1 & TPPtr_human_results$model_converged_treated_2 & TPPtr_human_results$model_converged_vehicle_1 & TPPtr_human_results$model_converged_vehicle_2]),
  "NPARC"=unique(NPARC_test_emp$id),
  "SCAM"=unique(DIM_Human_SCAM$Accession),"MSstatsTMT"=unique(DIM$ComparisonResult$Protein))
#remove shared accessions
Benchmarks_TPPproc<-lapply(Benchmarks_TPPproc,function(x) x[!stringr::str_detect(x,";")])
#remove missing values
Benchmarks_TPPproc<-lapply(Benchmarks_TPPproc,function(x) x[!is.na(x)])

#verify if targets are present in the Statistical model fit convergence
Benchmarks_TPPproc_targ<-purrr::map(Benchmarks_TPPproc,function(x) sum(x%in%c("P36507","Q02750")))

Benchmarks_TPPproc_n<-purrr::map2(Benchmarks_TPPproc,Benchmarks_TPPproc_targ,function(x,y) data.frame(Proteins=as.character(length(unique(x))),ver_targets=as.character(y)))|>dplyr::bind_rows()|>dplyr::mutate(Steps="Proteins: converged fit\n(Table 2, steps 1-3)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```

For proteins that passed the R\^2 and plateau filters

```{r}
TPPtr_filters<-unique(TPPtr_human_results$Protein_ID[TPPtr_human_results$passed_filter_vehicle_1_vs_treated_1&TPPtr_human_results$passed_filter_vehicle_2_vs_treated_2])
#remove shared accessions
TPPtr_filters<-TPPtr_filters[!stringr::str_detect(TPPtr_filters,";")]
#remove missing values
TPPtr_filters<-TPPtr_filters|>na.omit()


Benchmarks_TPPproc_filt<-purrr::map2(Benchmarks_TPPproc,names(Benchmarks_TPPproc),function(x,y) ifelse(y=="NPARC",sum(x %in% TPPtr_filters),sum(x%in%x)))

Benchmarks_TPPproc_filt<-lapply(Benchmarks_TPPproc_filt,function(x) as.character(x)|>na.omit())
#verify if targets are present in the filtered data
Benchmarks_TPPproc_targ<-
  purrr::map2(Benchmarks_TPPproc,names(Benchmarks_TPPproc),function(x,y) ifelse(y=="NPARC",sum(x %in% TPPtr_filters& x%in%c("P36507","Q02750")),sum(x%in%c("P36507","Q02750"))))

Benchmarks_TPPproc_filt_n<-purrr::map2(Benchmarks_TPPproc_filt,Benchmarks_TPPproc_targ,function(x,y) data.frame(Proteins=as.character(x),ver_targets=as.character(y)))|>dplyr::bind_rows()|>dplyr::mutate(Steps="Proteins: after filtering \n(Table 2, steps 1-4)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```

For differentially expressed proteins

```{r}
Benchmarks_TPPproc_p<-list("TPP"=unique(TPPtr_human_results$Protein_ID[TPPtr_human_results$pVal_adj_vehicle_1_vs_treated_1<0.05 & TPPtr_human_results$pVal_adj_vehicle_2_vs_treated_2 < 0.05 & TPPtr_human_results$Protein_ID %in% TPPtr_filters]),"NPARC"=as.character(unique(NPARC_test_emp$id[NPARC_test_emp$pAdj<0.05&NPARC_test_emp$id %in% TPPtr_filters])),"SCAM"=unique(DIM_Human_SCAM$Accession[DIM_Human_SCAM$ATE_padjBH<0.05]),"MSstatsTMT"=unique(DIM$ComparisonResult$Protein[DIM$ComparisonResult$adj.pvalue<0.05]))
#remove shared accessions
Benchmarks_TPPproc_p<-lapply(Benchmarks_TPPproc_p,function(x) x[!stringr::str_detect(x,";")])
#remove missing values
Benchmarks_TPPproc_p<-lapply(Benchmarks_TPPproc_p,function(x) as.character(x)|>na.omit())

#verify if differentially abundant proteins are present in the filtered data
Benchmarks_TPPproc_targ<-
  purrr::map2(Benchmarks_TPPproc_p,names(Benchmarks_TPPproc_p),function(x,y) ifelse(y=="NPARC",sum(x %in% TPPtr_filters&x%in%c("P36507","Q02750")),sum(x%in%c("P36507","Q02750"))))

Benchmarks_TPPproc_p_n<-purrr::map2(Benchmarks_TPPproc_p,Benchmarks_TPPproc_targ,function(x,y) data.frame(Proteins=as.character(length(unique(x))),ver_targets=as.character(y)))|>dplyr::bind_rows()|>dplyr::mutate(Steps="Proteins: differential abundance \n(Table 2, steps 1-5)",Analysis=c("TPP","NPARC","SCAM","MSstatsTMT"))
```

## Plot the bar charts

```{r}
bar_human<-dplyr::bind_rows(Total_prot_n,Output_proc_n,Benchmarks_TPPproc_n,Benchmarks_TPPproc_filt_n,Benchmarks_TPPproc_p_n)

bar_human$Analysis<-factor(bar_human$Analysis,levels=c("TPP","NPARC","SCAM","MSstatsTMT"))

col<-data.frame(Analysis=factor(unique(bar_human$Analysis),levels=c("TPP","NPARC","SCAM","MSstatsTMT")),"stat_model"=c("#D95F0E","#FEC44F","#2C7FB8","#07fff8"))


bar_human<-bar_human|>dplyr::inner_join(col)

bar_human$Analysis<-factor(paste0("TPP proc., ",bar_human$Analysis," model"),levels=c("TPP proc., TPP model","TPP proc., NPARC model","TPP proc., SCAM model","TPP proc., MSstatsTMT model"))


bar_human$Steps<-factor(bar_human$Steps,levels=c("Proteins: differential abundance \n(Table 2, steps 1-5)","Proteins: after filtering \n(Table 2, steps 1-4)","Proteins: converged fit\n(Table 2, steps 1-3)","Proteins: after \n TPP normalization (Table 1)","Proteins (unique protein groups):\n from PD"))


bar_human$Protein_targ<-paste0(bar_human$Proteins," ","(",bar_human$ver_targets,")")

bHuman<-ggplot(bar_human|>dplyr::group_by(Analysis),mapping=aes(x=Steps,y=as.integer(Proteins),alpha=Steps,fill=stat_model))+geom_col(stat_count=aes(x=Steps,y=Proteins),colour="black")+scale_y_continuous(limits=c(0,15000),breaks = c(0,6000,12000))+ylab("Number of proteins remaining after each step")+coord_flip()+facet_wrap(~Analysis,nrow=1, labeller = label_wrap_gen(width=15))+scale_fill_manual(values=c("#2C7FB8","#07fff8","#D95F0E","#FEC44F"))+
  xlab("Steps")+
  geom_text(aes(label=Protein_targ),size=8,position=position_dodge(width=0.5),hjust=-0.350,alpha=0.9)+theme(text=element_text(size=25,                          color = "black"),axis.text.x = element_text(size=25,face="bold"),axis.text.y=element_text(size=25,face="bold"),
                                                                                                            axis.title.x = element_text(size=25),axis.title.y=element_text(size=25),strip.text.x=element_text(size=25),legend.position="none", strip.placement = "outside")

path = getwd()
bHuman
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Bar_charts_TPP.tiff")

```

# Compare Venn Diagrams between processing (copy and paste in console)

```{r}


MSstatsTMT_proc<-lapply(Thermallate_Tm_MSstatsTMTprocesed
                        ,function(x) as.character(x)|>na.omit())

TPP_proc<-lapply(Thermallate_Tm_TPPprocesed,function(x) as.character(x)|>na.omit())


compare_proc<-purrr::map2(TPP_proc,MSstatsTMT_proc,function(x,y) list("MSstatsTMT proc."=y,"TPP proc."=x))


colors<-list(TPP=c("#D95F0E","#D95F0E"),NPARC=c("#FEC44F","#FEC44F"),SCAM=c("#07fff8","#07fff8"),MSstatsTMT=c("#2C7FB8","#2C7FB8"))

colors<-purrr::set_names(colors,names(colors))
purrr::map2(compare_proc[1],colors[1],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Human_processing ',"TPP"),
  output = TRUE ,
  imagetype="png" ,
  height = 600 , 
  width = 600 , 
  resolution = 1200,
  compression = "lzw",
  lwd = 0.5,
  col=y,
  cex=0.3,
  cat.cex=0.15,
  rotation.degree=-90,
     cat.pos = c(180,0),
          cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))
  

purrr::map2(compare_proc[2],colors[2],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Human_processing ',"NPARC"),
  output = TRUE ,
  imagetype="png" ,
  height = 600 , 
  width = 600 , 
  resolution = 1200,
  compression = "lzw",
  lwd = 0.5,
  col=y,
  cex=0.3,
  cat.cex=0.15,
  rotation.degree=90,
    cat.pos = c(180,0),
    cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))


purrr::map2(compare_proc[3],colors[3],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Human_processing ',"SCAM"),
  output = TRUE ,
  imagetype="png" ,
  height = 600 , 
  width = 600 , 
  resolution = 1200,
  compression = "lzw",
  lwd = 0.5,
  col=y,
  cex=0.3,
  cat.cex=0.15,
  rotation.degree=90,
    cat.pos = c(180,0),
    cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))

colors<-purrr::set_names(colors,names(colors))
purrr::map2(compare_proc[4],colors[4],function(x,y)VennDiagram::venn.diagram(
  x = x,
  category.names =names(x),
  filename = paste0('Venn_Human_processing ',"MSstatsTMT"),
  output = TRUE ,
  imagetype="png" ,
  height = 600 , 
  width = 600 , 
  resolution = 1200,
  compression = "lzw",
  lwd = 0.5,
  col=y,
  cex=0.3,
  cat.cex=0.15,
  rotation.degree=90,
     cat.pos = c(180,0),
          cat.dist = c(0.02, 0.02),
  scaled=FALSE,
  inverted=FALSE
  
))

```

#Save results for Venn diagram processing

## 

```{r}
names(Benchmarks_TPPproc_p)<-c("TPP stat. model","NPARC stat. model","SCAM stat. model","MSstatsTMT stat. model")

saveRDS(Benchmarks_TPPproc_p,"Human_TPPprocesed.RDS")
```

# Venn Diagrams

```{r}

Venn_Humandata<-ggvenn::ggvenn(data=Benchmarks_TPPproc_p,fill_color =c("#D95F0E","#FEC44F","#07fff8", "#2C7FB8"),set_name_size=4,text_size=4,stroke_size=1.1,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))
Venn_Humandata


```

# Venn Diagrams (top 3)

```{r}
png(filename = "Venn_Humandata_Tm_TPP.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
Venn_top3<-ggvenn::ggvenn(data=Benchmarks_TPPproc_p[c(1,2,4)],fill_color =c("#D95F0E","#FEC44F","#2C7FB8"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("TPP processed pAdj < 0.05")+theme(text=element_text(size=20))
Venn_top3
dev.off() 
```

#Filter expected targets from each result

```{r}
if(length(Benchmarks_TPPproc_p$`NPARC`)==0){
  NPARC_human<-data.frame(id=rep(NA,2),pAdj=rep(1,2))
}else{
  NPARC_human<-NPARC_test_emp|>dplyr::filter(id %in% c("P36507","Q02750"))
}
TPP_human<-TPP_Human_results_F|>dplyr::filter(uniqueID %in% c("P36507","Q02750"))

SCAM_human<-DIM_Human_SCAM|>dplyr::filter(Accession %in% c("P36507","Q02750"))

MSstat_human<-DIM$ComparisonResult|>dplyr::filter(Protein%in%c("P36507","Q02750"))


Stat_mod<-data.frame()
Stat_mod<-data.frame(Protein=unique(MSstat_human$Protein))|>dplyr::mutate(NPARC_adjpval=NPARC_human$pAdj,TPP_adjpval=TPP_human$p_adj_NPARC,SCAM_adjpval=SCAM_human$ATE_padjBH,MSstats_adjpval=MSstat_human$adj.pvalue)


```

# Plot the curves

```{r}
targets<-Human_TPPnorm_Proteins$normData|>as.data.frame()|>
  dplyr::filter(Protein%in% c("P36507","Q02750"))|>
  dplyr::mutate(Accession=Protein,
                treatment=stringr::str_extract(Experiment,"[[:lower:]]+"))|>
  dplyr::inner_join(Stat_mod,relationship="many-to-many")|>
  dplyr::group_by(Accession)|>
  dplyr::group_split()

DIM_targets<-DIM$FittedModel[which(names(DIM$FittedModel) %in% c("P36507","Q02750"))]
fitted_MSstat<-purrr::map2(targets,DIM_targets,function(x,y)x|>dplyr::mutate(model=list(y)))

nam<-as.character(unique(dplyr::bind_rows(targets)$Protein))
target_curves<-lapply(fitted_MSstat,function(x){plotMethods(x,temps=c("53.8","57.1","60.4"),processing="TPP",fit="Spline")
})
names(target_curves)<-nam

target_curves$P36507
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Curves_TPP_spline_P36507.tiff")

target_curves$Q02750
ggplot2::ggsave(path = path, width = 24, height = 6.5, device='tiff', dpi=600, filename = "Curves_TPP_spline_Q02750.tiff")

```

See which proteins are significant with MSsstatsTMT

```{r}
MSstats_sig<-as.character(DIM$ComparisonResult$Protein[which(DIM$ComparisonResult$adj.pvalue<0.05)])

MSstats_targets<-MSstats_sig[which(MSstats_sig %in% c("P36507","Q02750"))]
```

Check NPARC significant proteins

```{r}
NPARC_sig<-as.character(NPARC_test_emp$id[which(NPARC_test_emp$pAdj<0.05)])

NPARC_targets<-NPARC_sig[which(NPARC_sig %in% c("P36507","Q02750"))]
```

Check TPP significant proteins

```{r}
TPP_sig<-as.character(TPP_Human_results_F$uniqueID[which(TPP_Human_results_F$p_adj_NPARC<0.05)])

TPP_targets<-TPP_sig[which(TPP_sig %in% c("P36507","Q02750"))]
```

Check SCAM targets

```{r}
DIM_sig<-as.character(DIM_Human_SCAM$Accession[which(DIM_Human_SCAM$ATE_padjBH<0.05)])

DIM_targets<-DIM_sig[which(DIM_sig %in% c("P36507","Q02750"))]
```

Create a table summary

```{r}
data<-data.frame(Method=c("TPP","NPARC","SCAM","MSstatsTMT"),"pAdj less than 0.05"=c(length(TPP_sig),length(NPARC_sig),length(DIM_sig),length(MSstats_sig)),targetsFound=c(ifelse(length(TPP_targets)==0,0,length(TPP_targets)),ifelse(length(NPARC_targets)==0,0,length(NPARC_targets)),ifelse(length(DIM_targets)==0,0,length(DIM_targets)),ifelse(length(MSstats_targets)==0,0,length(MSstats_targets))))
data
```

# Volcano plot for MSstatsTMT

```{r}


f<-DIM$ComparisonResult
f$diffexpressed <- "Not Shifted"
    f$diffexpressed[f$adj.pvalue < 0.05] <- "Shifted adj. pVal < 0.05"
    f$diffexpressed[f$adj.pvalue < 0.01] <- "Shifted adj. pVal < 0.01"
    f$diffexpressed<-as.factor(f$diffexpressed)
    f$targets<-NA
    f$delabel <- NA
    f$others<-NA
    
    f$others[which(f$adj.pvalue<0.01)]<-as.character(f$Protein[which(f$adj.pvalue<0.01)])
    f$targets[f$Protein %in%c("Q7ZUY3","Q6NV46",                 "Q68SP3",# stat5a,
                                                                           "A4QNT9",# stat5b,
                                                                           "Q90Y03",# aldh1a2,
                                                                           "Q0H2G3",# aldh1a3,
                                                                           "F1QZU7",# aldh1a2.2,
                                                                           "A2BGR9",# aldh1a2.1,
                                                                           "F1QLV5",#nqo1,
                                                                           "F1Q7F3")] <- as.character(f$Protein[f$Protein %in%c("Q7ZUY3","Q6NV46",                 "Q68SP3",# stat5a,
                                                                           "A4QNT9",# stat5b,
                                                                           "Q90Y03",# aldh1a2,
                                                                           "Q0H2G3",# aldh1a3,
                                                                           "F1QZU7",# aldh1a2.2,
                                                                           "A2BGR9",# aldh1a2.1,
                                                                           "F1QLV5",#nqo1,
                                                                           "F1Q7F3")])
    f$delabel[f$Protein %in%c("P36507","Q02750","P40763","Q9Y2Q5","P31785")] <- as.character(f$Protein[f$Protein %in%c("P36507","Q02750","P40763","Q9Y2Q5","P31785")])
    flevels<-data.frame(colors=c("#762a83","#b35806","#1b7837"),
                        labels=c("Shifted adj. pVal < 0.05","Not Shifted","Shifted adj. pVal < 0.01"))
    flevels<-flevels %>% dplyr::filter(labels %in% unique(f$diffexpressed))

ggplot(f,mapping=aes(x=log2FC,y=log10(adj.pvalue),color=diffexpressed))+geom_point()+xlim(-4,4)+xlab("DIM")+scale_y_reverse()+geom_label_repel(f,mapping=aes(label=targets),color="black",max.overlaps = getOption("ggrepel.max.overlaps", default = 10),
nudge_x = 3,
nudge_y=1,
force = 3)+ylab("log(adj.pvalue)")+xlab("DIM")+theme(text=element_text(size=18))
```

# Volcano plot for SCAM

```{r}


f<-DIM$ComparisonResult
f$diffexpressed <- "Not Shifted"
    f$diffexpressed[f$adj.pvalue < 0.05] <- "Shifted adj. pVal < 0.05"
    f$diffexpressed[f$adj.pvalue < 0.01] <- "Shifted adj. pVal < 0.01"
    f$diffexpressed<-as.factor(f$diffexpressed)
    f$targets<-NA
    f$delabel <- NA
    f$others<-NA
    
    f$others[which(f$adj.pvalue<0.01)]<-as.character(f$Protein[which(f$adj.pvalue<0.01)])
    f$targets[f$Protein %in%c("Q7ZUY3","Q6NV46",                 "Q68SP3",# stat5a,
                                                                           "A4QNT9",# stat5b,
                                                                           "Q90Y03",# aldh1a2,
                                                                           "Q0H2G3",# aldh1a3,
                                                                           "F1QZU7",# aldh1a2.2,
                                                                           "A2BGR9",# aldh1a2.1,
                                                                           "F1QLV5",#nqo1,
                                                                           "F1Q7F3")] <- as.character(f$Protein[f$Protein %in%c("Q7ZUY3","Q6NV46",                 "Q68SP3",# stat5a,
                                                                           "A4QNT9",# stat5b,
                                                                           "Q90Y03",# aldh1a2,
                                                                           "Q0H2G3",# aldh1a3,
                                                                           "F1QZU7",# aldh1a2.2,
                                                                           "A2BGR9",# aldh1a2.1,
                                                                           "F1QLV5",#nqo1,
                                                                           "F1Q7F3")])
    f$delabel[f$Protein %in%c("P36507","Q02750","P40763","Q9Y2Q5","P31785")] <- as.character(f$Protein[f$Protein %in%c("P36507","Q02750","P40763","Q9Y2Q5","P31785")])
    flevels<-data.frame(colors=c("#762a83","#b35806","#1b7837"),
                        labels=c("Shifted adj. pVal < 0.05","Not Shifted","Shifted adj. pVal < 0.01"))
    flevels<-flevels %>% dplyr::filter(labels %in% unique(f$diffexpressed))

ggplot(f,mapping=aes(x=log2FC,y=log10(adj.pvalue),color=diffexpressed))+geom_point()+xlim(-4,4)+xlab("DIM")+scale_y_reverse()+geom_label_repel(f,mapping=aes(label=targets),color="black",max.overlaps = getOption("ggrepel.max.overlaps", default = 10),
nudge_x = 3,
nudge_y=1,
force = 3)+ylab("log(adj.pvalue)")+xlab("DIM")+theme(text=element_text(size=18))
```
