oad

---
title: "Data_Processing_MSstatsTMT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")

```

Library setup

```{r setup}

library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())

```

# Data Import

# Import option 1: Read Protein file from folder

```{r}
CETSA_OnePot<-readxl::read_xlsx("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/OnePotTPPStauro_Fx_Proteins.xlsx")
colnames(CETSA_OnePot) <- stringr::str_replace_all(colnames(CETSA_OnePot),"[:punct:]| ",".")
CETSA_OnePot_PSMs<-read.delim("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/OnePotTPPStauro_Fx_PSMs.txt")


```

# Format the data

```{r}
CETSA_OnePot<-tidyr::pivot_longer(CETSA_OnePot,cols=names(CETSA_OnePot)[stringr::str_detect(names(CETSA_OnePot),"Abundance")],names_to="Channel",values_to="Abundance")

CETSA_OnePot_PSMs<-tidyr::pivot_longer(CETSA_OnePot_PSMs,cols=names(CETSA_OnePot_PSMs)[stringr::str_detect(names(CETSA_OnePot_PSMs),"Abundance")],names_to="Channel",values_to="Abundance")
```

# Log transform

```{r}
CETSA_OnePot<-CETSA_OnePot|>dplyr::mutate(Abundance=log2(Abundance),color="#b11226",Channel=stringr::str_extract(Channel,"[:digit:][:digit:][:digit:]C|[:digit:][:digit:][:digit:]N|[:digit:][:digit:][:digit:]"))

CETSA_OnePot_PSMs<-CETSA_OnePot_PSMs|>dplyr::mutate(Abundance=log2(Abundance),color="gray",Channel=stringr::str_extract(Channel,"[:digit:][:digit:][:digit:]C|[:digit:][:digit:][:digit:]N|[:digit:][:digit:][:digit:]"))
```

# Annotation file

```{r}
CETSA_OnePot_annotation<-data.table::fread("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/OnePotTPP_annotation.csv")


```

```{r}
CETSA_OnePot$Protein<-CETSA_OnePot$Accession
CETSA_OnePot_PSMs$Protein<-CETSA_OnePot_PSMs$Master.Protein.Accessions

CETSA_OnePot_annotation$Condition<-stringr::str_remove(CETSA_OnePot_annotation$BioReplicate,"_rep[:digit:]")
class(CETSA_OnePot_annotation$TechRepMixture)<-"numeric"
CETSA_OnePot_annotation$TechRepMixture<-1
CETSA_OnePot_annotation$Mixture<-1

CETSA_OnePot_annotation<-CETSA_OnePot_annotation|>dplyr::select(-Fraction)|>dplyr::distinct()
CETSA_OnePot_annotation$Run<-"OnePot"
CETSA_OnePot_annotation<-CETSA_OnePot_annotation|>dplyr::distinct()

#rename condition
CETSA_OnePot_annotation<-CETSA_OnePot_annotation|>as.data.frame()|>dplyr::group_split(Condition)
CETSA_OnePot_annotation<-lapply(CETSA_OnePot_annotation,function(x){
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
  return(x)})
CETSA_OnePot_annotation<-dplyr::bind_rows(CETSA_OnePot_annotation)
#right join annotation to data
CETSA_OnePot<-CETSA_OnePot|>dplyr::inner_join(CETSA_OnePot_annotation)
CETSA_OnePot_PSMs<-CETSA_OnePot_PSMs|>dplyr::inner_join(CETSA_OnePot_annotation)
```

# Plot the profile plots

```{r}
MEK2<-CETSA_OnePot|>dplyr::filter(Protein=="P36507")|>dplyr::mutate(Channel=factor(Channel))

levels(MEK2$Channel)<-set_temps(16,seq(1,16))$Channel[seq(1,length(unique(CETSA_OnePot$Channel)))]

MEK2<-MEK2|>dplyr::arrange(Channel)

MEK2$order<-seq(1,15)

ggplot(MEK2,mapping=aes(x=order,y=Abundance,color=MEK2$color[1]))+geom_point()+geom_line()+ylim(0,12.5)+theme_set(theme_bw())+ylab(expression(log[2]~Abundance))+theme(axis.title.x = element_blank(),axis.text.y = element_text(size=20),axis.title.y = element_text(size=20))+geom_vline(mapping=aes(x=3.5))
```

# Format original data to MSstatsTMT

```{r}
WD<-"~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/"
PD_df<-read_cetsa_concentrations(WD,WD,Frac=TRUE,concentrations=set_concentration(16,concentrations=c(unique(CETSA_OnePot_annotation$Condition),NA),replicates=3),solvent="Control",CARRIER=FALSE)


#Convert to msstatsTMT format

PD_df<-PD_df|>dplyr::mutate(Protein=as.character(Accession),
                            Run="OnePot",
                            Abundance=log2(Abundance))|>
  dplyr::select(Protein,Run,Abundance,Channel,Gene.Symbol)|>
  dplyr::distinct()|>
  dplyr::inner_join(CETSA_OnePot_annotation)
                            
MSstats_input<-list(ProteinLevelData=PD_df,FeatureLevelData=NA)                           
```

# Normalization with TPP CCR requires more than one plex

```{r}
data(hdacCCR_smallExample)
tppccrData <- tppccrImport(configTable=hdacCCR_config, data = hdacCCR_data)
head(hdacCCR_data$Panobinostat_1)
```

# GroupComparisonTMT DMSO OnePot vs 25x

```{r}
contrast.matrix<-matrix(ncol=5,nrow=1)
colnames(contrast.matrix)<-unique(PD_df$Condition)
contrast.matrix[1,]<-c(-1,0,0,0,1)
rownames(contrast.matrix)<-c("DMSOvs25x")
DIM_DMSOvs25x<-MSstatsTMT::groupComparisonTMT(MSstats_input,
                   contrast.matrix=contrast.matrix,
                   moderated=FALSE,
                   adj.method="BH",
                   remove_norm_channel = FALSE,
                   remove_empty_channel = FALSE,
                   save_fitted_models = TRUE,
                   use_log_file = FALSE,
                   append=FALSE)
```

Determine overlaps between MSstatsTMT and OnePot

```{r}
#read in kinase file
WD<-"~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()
setwd(WD)
Proteins<-data.frame()
Proteins<-lapply(h,function(x) readxl::read_xlsx(paste0(WD,x))) |> dplyr::bind_rows()|>dplyr::select(Accession,"Gene Symbol")|>dplyr::rename("GeneSymbol"="Gene Symbol")|>dplyr::distinct()

kinases<-readxl::read_xlsx(paste0(WD,"/Kinase accessions.xlsx"))|>dplyr::inner_join(Proteins,relationship="many-to-many")

kinase_interactors<-read.delim(paste0(WD,"kinase_plus_direct_interactions.txt"))|>dplyr::mutate(GeneSymbol=Gene)

kinase_df<-PD_df|>dplyr::mutate(Accession=Protein)|>dplyr::inner_join(kinases,by="Accession",relationship="many-to-many")
#join geneSymbols to original DF
kinase_df<-kinase_df|>dplyr::inner_join(Proteins,relationship="many-to-many")
kinases<-kinase_df$Accession[kinase_df$GeneSymbol %in% kinase_interactors$GeneSymbol]
kinases<-unique(kinases)

#Published work
OnePotTPP_2D_hits_kinases<-unique(kinase_df$Protein[stringr::str_detect(kinase_df$`DetectedOnePotTPP?`,"Y")&kinase_df$Protein%in% kinases])|>na.omit()

SPROX_2D_hits_kinases<-unique(kinase_df$Protein[stringr::str_detect(kinase_df$`DetectedOnePotSPROX?`,"Y")&kinase_df$Protein%in% kinases])|>na.omit()

OnePot_SPROX_hits_kinases<-unique(kinase_df$Protein[stringr::str_detect(kinase_df$`OnePotSPROXHits?`,"Y")&kinase_df$Protein%in% kinases])|>na.omit()

TPP_hits_kinases<-unique(kinase_df$Protein[stringr::str_detect(kinase_df$`OnePotTPPHits?`,"Y")&kinase_df$Protein%in% kinases])|>na.omit()



```

NPARC

```{r}

Human_TPP_TPPprocesed <- readRDS("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/Reprocessed/Human_TPP_TPPprocesed.RDS")

names(Human_TPP_TPPprocesed)<-c("NPARC temp.","TPP temp.","SCAM temp.","MSstatsTMT temp.")


MSstat_hits<-unique(stringr::str_remove(as.character(DIM_DMSOvs25x$ComparisonResult$Protein[DIM_DMSOvs25x$ComparisonResult$adj.pvalue<0.05]),"-[:digit:]"))
MSstat_hits<-MSstat_hits[!stringr::str_detect(MSstat_hits,";")]
#add benchmarks to original full curve designs
Benchmarks_external<-append(list("NPARC temp."=Human_TPP_TPPprocesed$`NPARC temp.`),list(
  "STRING targets"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennNPARC_Fulltemp_vs_OnePot_TPPproc.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#FEC44F","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 

```

TPP

```{r}

#add benchmarks to original full curve designs
Benchmarks_external<-append(list("TPP temp."=Human_TPP_TPPprocesed$`TPP temp.`),list(
  "STRING targets"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennTPP_Fulltemp_vs_OnePot_TPPproc.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#D95F0E","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 

```

SCAM

```{r}
Benchmarks_external<-append(list("SCAM temp."=Human_TPP_TPPprocesed$`SCAM temp.`),list(
  "STRING targets"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennSCAM_Fulltemp_vs_OnePot_MSstatsTMTproc.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#07fff8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 
```

MSstatsTMT

```{r}
Benchmarks_external<-append(list("MSstatsTMT temp."=Human_TPP_TPPprocesed$`MSstatsTMT temp.`),list(
  "STRING targets"=na.omit(kinases),
                         "MSstatsTMT OnePot DMSO vs 25x"=MSstat_hits))

png(filename = "VennMSstatsTMT_Fulltemp_vs_OnePot_MSstatsTMTproc.png",
    width =600, height = 600, units = "px", pointsize = 12,
    res = 130,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#2c7fb8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+ggtitle("pAdj < 0.05")+theme(text=element_text(size=20))

dev.off() 

saveRDS(MSstat_hits,"PD_Proteinproc_OnePot.RDS")
```

# Profile plots

```{r}
MEK2<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="P36507")|>dplyr::group_split(Condition)
MEK2_renamed<-lapply(MEK2,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
ggplot(MEK2_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_point()+ylab(expression(log[2]~Abundance))

```

box plots

```{r}
MEK2<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="P36507")|>dplyr::group_split(Condition)
MEK2_renamed<-lapply(MEK2,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
MEK2_plotOnePot<-ggplot(MEK2_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_boxplot()+ylab(expression(log[2]~Abundance))+ggtitle("MAP2K2")
saveRDS(MEK2_plotOnePot,"MEK2_allMethods_sig.RDS")
MEK2_plotOnePot
```

Boxplots for Syntaxin 2

```{r}
STX2<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="P32856")|>dplyr::group_split(Condition)
STX2_renamed<-lapply(STX2,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
STX2_plotOnePot<-ggplot(STX2_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_boxplot()+ylab(expression(log[2]~Abundance))+ggtitle("STX2")

saveRDS(STX2_plotOnePot,"STX2_MSstats_sigOnePot.RDS")
STX2_plotOnePot
```

```{r}
CHK1<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="O14757")|>dplyr::group_split(Condition)
CHK1_renamed<-lapply(CHK1,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
CHK1_plotOnePot<-ggplot(CHK1_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_boxplot()+ylab(expression(log[2]~Abundance))+ggtitle("CHK1")

saveRDS(CHK1_plotOnePot,"STX2_MSstats_sigOnePot.RDS")
CHK1_plotOnePot
```
