------------------------------------------------------------------------

---
title: "Vignette8"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#With dataset 3b, this vignette will outline OnePot processed datasets at the peptide and evaluate MSstatsTMT statistical modeling.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")

```

Library setup

```{r setup}

library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())

```

#A. Data Import

### Import option 1: Read in your own file

**Use getwd(your_working_directory) to get a folder where your PSM data is stored**

```{r}

# h<-list.files(WD,pattern="PSMs")|>as.list()
# 
# df<-purrr::map(h,function(x) readxl::read_xlsx(path=paste0(WD,"/",x),.name_repair = "universal")) |> dplyr::bind_rows()
# head(df)

```

### Import option 2: Read PSM file from folder

```{r}
CETSA_OnePot<-data.table::fread("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/OnePotTPPStauro_Fx_PSMs.txt")
colnames(CETSA_OnePot) <- stringr::str_replace_all(colnames(CETSA_OnePot),"[:punct:]| ",".")

#save original data
df<-CETSA_OnePot
head(CETSA_OnePot)
```

#B. Annotation File Import
Read annotation file

```{r}
CETSA_OnePot_annotation<-data.table::fread("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/OnePotTPP_annotation.csv")

class(CETSA_OnePot_annotation$TechRepMixture)<-"numeric"
CETSA_OnePot_annotation$TechRepMixture<-1
CETSA_OnePot_annotation$Mixture<-1
CETSA_OnePot_annotation$Condition<-stringr::str_remove(CETSA_OnePot_annotation$BioReplicate,"_rep[:digit:]")

```

#C. Data processing with MSstatsTMT

Import the data into MSStatsTMT

```{r}


processed.features <- MSstatsTMT::PDtoMSstatsTMTFormat(CETSA_OnePot,
                                                    CETSA_OnePot_annotation,
                                                    which.proteinid = "Master.Protein.Accessions",
                                                    useNumProteinsColumn = FALSE,
                                                    useUniquePeptide = TRUE,
                                                    rmPSM_withfewMea_withinRun = TRUE,
                                                    rmProtein_with1Feature = TRUE,
                                                    summaryforMultipleRows = max,
                                                    use_log_file = FALSE,
                                                    append=FALSE,
                                                    verbose=TRUE)
```
#D. Summarize the peptide data into proteins

```{r}

                                                   
summarised.proteins <- MSstatsTMT::proteinSummarization(data = processed.features,
                                method = "Median",
                                global_norm = FALSE,
                                reference_norm = FALSE,
                                MBimpute=TRUE,
                                use_log_file=FALSE,
                                append=FALSE,
                                remove_norm_channel=FALSE,
                                remove_empty_channel=FALSE)
```
#E. Visualize a profile plot for MEK2

    ```{r}

    MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P36507")
    ```

#F. Fit linear mixed models with groupComparisons from MSstatsTMT

```{r}
DIM_OnePot<-groupComparisonTMT(
  summarised.proteins,
  contrast.matrix = "pairwise",
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)

```
#G. Compare DMSO vs 25x staurosporine

```{r}
contrast.matrix<-matrix(nrow=1,ncol=5)
colnames(contrast.matrix)<-unique(df3$ProteinLevelData$Condition)
contrast.matrix[1,]<-c(-1,0,0,0,1)
rownames(contrast.matrix)<-"DMSOvs25x"
DIM_OnePot_DMSOvs25x<-groupComparisonTMT(
  df3,
  contrast.matrix = contrast.matrix,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL
)
```
#H. Compare DMSO vs all concentrations on avg

```{r}
contrast.matrix<-matrix(nrow=1,ncol=5)
colnames(contrast.matrix)<-unique(df3$ProteinLevelData$Condition)
contrast.matrix[1,]<-c(-1,1/4,1/4,1/4,1/4)
rownames(contrast.matrix)<-"DMSOvsAllAvg"
DIM_OnePot_DMSOvsAllAvg<-groupComparisonThermalProfiling(
  df3,
  contrast.matrix = contrast.matrix,
  moderated = FALSE,
  adj.method = "BH",
  remove_norm_channel = TRUE,
  remove_empty_channel = TRUE,
  save_fitted_models = TRUE,
  use_log_file = FALSE,
  append = FALSE,
  verbose = TRUE,
  log_file_path = NULL,
  missing_timepoint = "replace",
  replacement=list("131_vehicle" = c("130C_vehicle"))
)
```
#I. Read in Human Kinome accessions and remove isoforms
```{r}
WD<-"~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()
setwd(WD)
Proteins<-data.frame()
Proteins<-lapply(h,function(x) readxl::read_xlsx(paste0(WD,x))) |> dplyr::bind_rows()|>dplyr::select(Accession,"Gene Symbol")|>dplyr::rename("GeneSymbol"="Gene Symbol")|>dplyr::distinct()
#Read in human kinome (522 proteins)
kinases<-read.csv(paste0(WD,"/HumanKinome_040920.csv"))

#remove isoforms from UniProt accessions
df<-df|>dplyr::mutate(Accession=stringr::str_remove(Master.Protein.Accessions,"-[[:digit:]]+"))|>dplyr::distinct()


#join Uniprot IDs to original DF
kinase_df<-kinases|>dplyr::inner_join(df,relationship="many-to-many")|>dplyr::select(Accession) |> dplyr::distinct()

```
#J. MSstats interactors

```{r}
#collect interactions from MSstatsTMT

MSstat_hits<-unique(DIM_OnePot_DMSOvs25x$ComparisonResult$Protein[DIM_OnePot_DMSOvs25x$ComparisonResult$adj.pvalue<0.05])
#remove shared protein accessions
MSstat_hits<-MSstat_hits[!stringr::str_detect(MSstat_hits,";")]

```

#K. Figure 4, panel E,  Read in MSstatsTMT processed results with TPP statistical models to compare against OnePot


```{r}

#add benchmarks to original full curve designs
Benchmarks_external<-append(list("3a: MSstatsTMT proc. 
TPP stat.mod."=Thermal_lateMSstatsTMTprocesed$`TPP temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: MSstatsTMT proc. 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennTPP_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#D95F0E","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 
```

#K. Figure 4, panel F,  Read in MSstatsTMT processed results with NPARC statistical models to compare against OnePot

```{r}
names(Thermal_lateMSstatsTMTprocesed)<-c("TPP temp.","NPARC temp.","SCAM temp.","MSstatsTMT temp.")
Thermal_lateMSstatsTMTprocesesed<-lapply(Thermal_lateMSstatsTMTprocesed, function(x) stringr::str_remove(x,"-[:digit:]")|>unique())
MSstat_hits<-unique(stringr::str_remove(as.character(DIM_OnePot_DMSOvs25x$ComparisonResult$Protein[DIM_OnePot_DMSOvs25x$ComparisonResult$adj.pvalue<0.05]),"-[:digit:]"))
#remove shared protein accessions
MSstat_hits<-MSstat_hits[!stringr::str_detect(MSstat_hits,";")]
#add benchmarks to original full curve designs
Benchmarks_external<-append(list("3a: MSstatsTMT proc. 
NPARC stat. mod."=Thermal_lateMSstatsTMTprocesed$'NPARC temp.'),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: MSstatsTMT proc. 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennNPARC_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
     width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")
library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#FEC44F","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))
dev.off() 

```
#L. Figure 4, panel G,  Read in MSstatsTMT processed results with SCAM statistical models to compare against OnePot
```{r}
#add benchmarks to original full curve designs
Benchmarks_external<-append(list("3a: MSstatsTMT proc. 
SCAM stat. mod."=Thermal_lateMSstatsTMTprocesed$`SCAM temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: MSstatsTMT proc. 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))
 

png(filename = "VennSCAM_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#07fff8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 
```
#M. Figure 4, panel H,  Read in MSstatsTMT processed results with MSstatsTMT statistical models to compare against OnePot

```{r}
#add benchmarks to original full curve designs
Benchmarks_external<-append(list("3a: MSstatsTMT proc. 
MSstatsTMT stat. mod."=Thermal_lateMSstatsTMTprocesed$`MSstatsTMT temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: MSstatsTMT proc. 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennMSstatsTMT_Fulltemp_vs_OnePot_MSstatsTMTproc_unique_Median.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#2c7fb8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 
```

#N. Figure 8, panel E: profile plots for MEK2

```{r}
#Rename columns for Conditions

summarised.proteins<-lapply(summarised.proteins,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","A_DMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","B_1xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","C_5xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","D_10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","E_25xstauro",x$Condition)
                            return(x)})
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P36507")
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="Q9Y5S2")
```

# Plot some targets not picked up by protein-level processing

```{r}
#sig only in TPP
 MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P32856")
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="Q9Y605")

#sig only in MSstatsTMT
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="Q9UBS4")
MSstatsTMT::dataProcessPlotsTMT(summarised.proteins,type="PROFILEPLOT",which.Protein="P16083")
```
#I. Figure 8, panel A: Venn Diagram for MSstatsTMT compared to the reference set and PD protein summary
```{r}

 
Benchmarks_external<-append(list("3a: TPP proc.
MSstatsTMT stat. mod."=MSstatsTMTstatmodel_TPPprocesed$`MSstatsTMT stat. model`),list(
 "Reference Set"=na.omit(kinases$Accession),
                         "3a MSstatsTMT proc.
 MSstatsTMT stat. mod."=MSstatsTMTproc.MSstatsTMTstatmodel))

png(filename = "VennMSstatsTMT_Fulltemp_vs_OnePot_PD_levelproc_onePot_unique_Median.png",
    width =12, height = 6, units = "in", pointsize = 14,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#1d547a","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 
```
#J.Figure 8, panel D: Venn Diagram for MSstatsTMT compared to the reference set and PD protein summary
```{r}

 
Benchmarks_external<-append(list("3b: PD protein summary
MSstatsTMT stat. mod."=PD_Proteinproc_OnePot),list(
 "Reference Set"=na.omit(kinases$Accession),
                         "3a MSstatsTMT proc.
 OnePot MSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennMSstatsTMT_OnePot_PD_levelproc_onePot_unique_Median.png",
    width =12, height = 6, units = "in", pointsize = 14,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#1d547a","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 
```
