
---
title: "Vignette7"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data_Processing_MSstatsTMT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include=TRUE}
devtools::load_all(".")

```

Library setup

```{r setup}

library(MSstats)
library(MSstatsConvert)
library(MSstatsTMT)
library(tidyr)
library(stringr)
library(TPP)
library(NPARC)
library(ggplot2)
library(ggrepel)
theme_set(theme_bw())

```
#With dataset 3b, this vignette will outline OnePot processed datasets at the protein level and evaluate TPP, NPARC, SCAM and MSstatsTMT statistical modeling.

#A. Data Import

### Import option 1: Read Protein file from folder

```{r}
CETSA_OnePot<-readxl::read_xlsx("D:/OnePot/Proteins.xlsx")
colnames(CETSA_OnePot) <- stringr::str_replace_all(colnames(CETSA_OnePot),"[:punct:]| ",".")
CETSA_OnePot_PSMs<-read.delim("D:/OnePot/PSMs.txt")
colnames(CETSA_OnePot) <- stringr::str_replace_all(colnames(CETSA_OnePot),"[:punct:]| ",".")
#save original data
df<-CETSA_OnePot
head(CETSA_OnePot)
```


#B. Annotation File import option

Read annotation file

```{r}
CETSA_OnePot_annotation<-data.table::fread("D:/OnePot/OnePotTPP_annotation.csv")


```

### Or generate annotation columns

```{r}
CETSA_OnePot$ProteinName<-CETSA_OnePot$Accession
CETSA_OnePot$Protein<-CETSA_OnePot$Accession
CETSA_OnePot_annotation$Condition<-stringr::str_remove(CETSA_OnePot_annotation$BioReplicate,"_rep[:digit:]")
class(CETSA_OnePot_annotation$TechRepMixture)<-"numeric"
CETSA_OnePot_annotation$TechRepMixture<-1
CETSA_OnePot_annotation$Mixture<-1

CETSA_OnePot_annotation<-CETSA_OnePot_annotation|>dplyr::distinct()
CETSA_OnePot_annotation$Run<-"OnePot"
CETSA_OnePot_annotation<-CETSA_OnePot_annotation|>dplyr::select(-Fraction)|>dplyr::distinct()
head(CETSA_OnePot)
```


#C. Format original PD protein data to MSstatsTMT

```{r}
WD<-"~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/CETSA_onePot/OnePot/"
PD_df<-read_cetsa_concentrations(WD,WD,Frac=TRUE,concentrations=set_concentration(16,concentrations=c(unique(CETSA_OnePot_annotation$Condition),NA),replicates=3),solvent="Control",CARRIER=FALSE)

#Convert to msstatsTMT format

PD_df<-PD_df|>dplyr::mutate(Protein=as.character(Accession),
                            Run="OnePot",
                            Abundance=log2(Abundance))|>
  dplyr::select(Protein,Run,Abundance,Channel,Gene.Symbol)|>
  dplyr::distinct()|>
  dplyr::inner_join(CETSA_OnePot_annotation)
                            
MSstats_input<-list(ProteinLevelData=PD_df,FeatureLevelData=NA)                           
```


#D. GroupComparisonTMT DMSO OnePot vs 25x as in Figure 8, Panel D for PD protein summary and panel E

```{r}
contrast.matrix<-matrix(ncol=5,nrow=1)
colnames(contrast.matrix)<-unique(PD_df$Condition)
contrast.matrix[1,]<-c(-1,0,0,0,1)
rownames(contrast.matrix)<-c("DMSOvs25x")
DIM_DMSOvs25x<-MSstatsTMT::groupComparisonTMT(MSstats_input,
                   contrast.matrix=contrast.matrix,
                   moderated=FALSE,
                   adj.method="BH",
                   remove_norm_channel = FALSE,
                   remove_empty_channel = FALSE,
                   save_fitted_models = TRUE,
                   use_log_file = FALSE,
                   append=FALSE)
```

#E. Read in protein data, human kinome and remove isoforms

```{r}
#read in protein file
WD<-"D:/OnePot/"
h<-list.files(WD,pattern="Proteins.xlsx")|>as.list()
setwd(WD)
Proteins<-data.frame()
Proteins<-lapply(h,function(x) readxl::read_xlsx(paste0(WD,x))) |> dplyr::bind_rows()|>dplyr::select(Accession,"Gene Symbol")|>dplyr::rename("GeneSymbol"="Gene Symbol")|>dplyr::distinct()
#Read in human kinome (522 proteins)
kinases<-read.csv(paste0(WD,"/HumanKinome_040920.csv"))
#remove isoforms from UniProt accessions
Proteins<-Proteins|>dplyr::mutate(Accession=stringr::str_remove(Accession,"-[[:digit:]]+"))|>dplyr::distinct()


#join Uniprot IDs to original DF
kinase_df<-kinases|>dplyr::inner_join(Proteins,relationship="many-to-many")|>dplyr::distinct()



```
#F. Figure 7, panel A,  Read in TPP processed PD results with TPP statistical models to compare against OnePot

```{r}

#add benchmarks to original full curve designs
Benchmarks_external<-append(list("3a: TPP proc. 
TPP stat. mod."=TPPcurve_TPPprocesed$`TPP temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: PD protein summary 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

grDevices::png(filename = "VennTPP_Fulltemp_vs_OnePot_TPP.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#D95F0E","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 

```

#F. Figure 7, panel B: Read in TPP processed PD results with NPARC statistical models to compare against OnePot

```{r}

TPPcurve_TPPprocesed <-readRDS("~/work/ivanovlab/figueroa-navedo.a/OnePot_dataset/Reprocessed/Human_TPP_TPPprocesed.RDS")

names(TPPcurve_TPPprocesed)<-c("TPP temp.","NPARC temp.","SCAM temp.","MSstatsTMT temp.")


MSstat_hits<-unique(stringr::str_remove(as.character(DIM_DMSOvs25x$ComparisonResult$Protein[DIM_DMSOvs25x$ComparisonResult$adj.pvalue<0.05]),"-[:digit:]"))
#remove shared protein accessions
MSstat_hits<-MSstat_hits[!stringr::str_detect(MSstat_hits,";")]
#add benchmarks to original full curve designs
Benchmarks_external<-append(list("3a: TPP proc. 
NPARC stat. mod."=TPPcurve_TPPprocesed$`NPARC temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: PD protein summary 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennNPARC_Fulltemp_vs_OnePot_TPPproc.png",
     width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")
library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#FEC44F","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))
dev.off() 

```


#G. Figure 7 panel C, Read in TPP processed PD results with SCAM statistical models to compare against OnePot

```{r}
Benchmarks_external<-append(list("3a: TPP proc.
SCAM stat. mod."=TPPcurve_TPPprocesed$`SCAM temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: PD protein summary 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennSCAM_Fulltemp_vs_OnePot_TPPproc.png", width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#07fff8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 
```

#H. Figure 7, panel D: Read in TPP processed PD results with MSstatsTMT statistical models to compare against OnePot

```{r}
Benchmarks_external<-append(list("3a: TPP proc. 
MSstatsTMT stat. mod."=TPPcurve_TPPprocesed$`MSstatsTMT temp.`),list(
  "Reference set"=na.omit(kinases$Accession),
                         "3b: PD protein summary 
  OnePotMSstatsTMT stat. mod."=MSstat_hits))

png(filename = "VennMSstatsTMT_Fulltemp_vs_OnePot_TPPproc.png",
    width =12, height = 6, units = "in", pointsize = 12,
    res = 600,type ="cairo")  

library(ggvenn)
ggvenn::ggvenn(data=Benchmarks_external,fill_color =c("#2c7fb8","#7b057b","#1d547a"),set_name_size=5,text_size=8,stroke_size=1.5,fill_alpha = 0.6,show_percentage = FALSE)+theme(text=element_text(size=20))+coord_cartesian(ylim = c(-2,2),xlim = c(-4,4))

dev.off() 

saveRDS(MSstat_hits,"PD_Proteinproc_OnePot.RDS")
```

#I. Figure 8, panel E: profile plots for MEK2

```{r}
MEK2<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="P36507")|>dplyr::group_split(Condition)
MEK2_renamed<-lapply(MEK2,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
ggplot(MEK2_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_point()+ylab(expression(log[2]~Abundance))

```

#J. Figure S-28, panel A: box plots for MEK2 

```{r}
MEK2<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="P36507")|>dplyr::group_split(Condition)
MEK2_renamed<-lapply(MEK2,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
MEK2_plotOnePot<-ggplot(MEK2_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_boxplot()+ylab(expression(log[2]~Abundance))+ggtitle("MAP2K2")
saveRDS(MEK2_plotOnePot,"MEK2_allMethods_sig.RDS")
MEK2_plotOnePot
```

#K. Figure S-28, panel B: box plots for Syntaxin 2

```{r}
STX2<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="P32856")|>dplyr::group_split(Condition)
STX2_renamed<-lapply(STX2,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
STX2_plotOnePot<-ggplot(STX2_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_boxplot()+ylab(expression(log[2]~Abundance))+ggtitle("STX2")

saveRDS(STX2_plotOnePot,"STX2_MSstats_sigOnePot.RDS")
STX2_plotOnePot
```
#L. Figure S-28, panel C: box plots for CHEK1
```{r}
CHEK1<-MSstats_input$ProteinLevelData|>dplyr::filter(Protein=="O14757")|>dplyr::group_split(Condition)
CHEK1_renamed<-lapply(CHEK1,function(x) {
  x$Condition<-as.character(x$Condition)
  x$Condition<-ifelse(x$Condition=="DMSO","00xDMSO",x$Condition)
  x$Condition<-ifelse(x$Condition=="1xstauro","01xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="5xstauro","05xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="10xstauro","10xstauro",x$Condition)
  x$Condition<-ifelse(x$Condition=="25xstauro","25xstauro",x$Condition)
                            return(x)})|>dplyr::bind_rows()
CHEK1_plotOnePot<-ggplot(CHEK1_renamed,mapping=aes(x=Condition,y=Abundance,color=Condition))+geom_boxplot()+ylab(expression(log[2]~Abundance))+ggtitle("CHK1")

saveRDS(CHEK1_plotOnePot,"CHK1_MSstats_sigOnePot.RDS")
CHEK1_plotOnePot
```
