---
title: "Simulation_dataset3b_OnePot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation_dataset3b_OnePot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#This vignette outlines a thermal pool design simulation for figure 6
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

library(ggplot2)
library(tidyr)

```

#A. Load in MSstatsTMT processed data

```{r}
#read summarised protein output from MSstatsTMT
summarised.proteins<- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation_data/summarised.proteinsD1.RDS")
summarised.proteinsNS<- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation_data/summarised.proteinsZebra.RDS")
```

#B.Figure 6, panel D design: 2 TMT 10-plexes, 10 biological replicates and 2 temperatures 
###Print out ICC distributions and simulate a strong interactor for a design with two temperatures, 2 plexes, 10 replicates

```{r}
#get variance components
Dataset_1_ICC<-msstats_icc(summarised.proteins,temps=unique(summarised.proteins$ProteinLevelData$temperature))
#generate simulations
 OnePot_2plex_10repsstrong<-benchmark_shifter_sim_2plex(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),n_replicates_per_plex =5,design="OnePot",shifter="strong")
Histograms_2temp_strong_onePot_10reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_10repsstrong,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex =5,plexes=2)

Histogram_strongOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_strongOnePot5_MsstatsTMTproc.RDS")
Histogram_strong10<-Histogram_strongOnePot_MsstatsTMTproc+theme(text=element_text(size=24))
path<-getwd()
Histogram_strong10
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_strong10.png")
```
###Simulate a weak interactor for a design with two temperatures, 2 plexes, 10 replicates per plex  by Condition

```{r}
#generate simulations
OnePot_2plex_10repsweak<-benchmark_shifter_sim_2plex(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="OnePot",shifter="weak")
Histograms_2temp_weak_onePot_10reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_10repsweak,design="OnePot",shifter="weak",t_range=seq(5,6),n_replicates_per_plex =5,variation_idx=NA,plexes=2)

Histogram_weakOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_weakOnePot5_MsstatsTMTproc.RDS")
Histogram_weak10<-Histogram_weakOnePot_MsstatsTMTproc+theme(text=element_text(size=24))
path<-getwd()
Histogram_weak10
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_weak10.png")
```
##Print out ICC values and simulate a non-interactor for a design with two temperatures, 2 plexes, 10 replicates per plex by Condition

```{r}
#get variance components
Dataset_2_ICC<-msstats_icc(summarised.proteinsNS,temps=unique(summarised.proteinsNS$ProteinLevelData$temperature))
#generate simulations

 OnePot_2plex_10repsnon<-benchmark_nonshifter_sim_2plex(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),n_replicates_per_plex =5,design="OnePot")
Histograms_2temp_non_onePot_10reps<-plot_benchmarks_MSstatsTMT_reps(  OnePot_2plex_10repsnon,design="OnePot",shifter="non",t_range=seq(5,6),n_replicates_per_plex =5,plexes=2)

Histogram_nonOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_nonOnePot5_MsstatsTMTproc.RDS")
Histogram_non10<-Histogram_nonOnePot_MsstatsTMTproc+theme(text=element_text(size=24))
path<-getwd()
Histogram_non10
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_non10.png")
```
#C. Figure 6, panel E: for a design with 1 TMT 10-plex, 1 biological replicate per mixture, 2 temperatures per plex and 2 temperatures as contrast
###Simulate a strong interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition
```{r}
OnePot_1plex_5repsstrong<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),design="OnePot",n_replicates_per_plex=5)
Histograms_2temp_strong_onePot_5reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_1plex_5repsstrong,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex =5,variation_idx = NA,plexes=1)
Histogram_strongOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_strongOnePot5_MsstatsTMTproc.RDS")
Histogram_strong5<-Histogram_strongOnePot_MsstatsTMTproc+theme(text=element_text(size=24))
path<-getwd()
Histogram_strong5
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_strong5.png")
```
###Simulate a weak interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition

```{r}
OnePot_1plex_5repsweak<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="OnePot",n_replicates_per_plex=5)

Histograms_2temp_weak_onePot_5reps<-plot_benchmarks_MSstatsTMT_reps(OnePot_1plex_5repsweak,design="OnePot",shifter="weak",t_range=seq(5,6),n_replicates_per_plex=5,variation_idx = NA,plexes=1)

Histogram_weakOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_weakOnePot5_MsstatsTMTproc.RDS")
Histogram_weak5<-Histogram_weakOnePot_MsstatsTMTproc+theme(text=element_text(size=24))
path<-getwd()
Histogram_weak5
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_weak5.png")
```

###Simulate a non interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition

```{r}
OnePot_1plex_5repsnon<-benchmark_nonshifter_sim_1plex(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),design="OnePot",n_replicates_per_plex=5)
Histograms_1temp_non_onePot_5reps<-plot_benchmarks_MSstatsTMT_reps(OnePot_1plex_5repsnon,design="OnePot",shifter="non",t_range=seq(5,6),n_replicates_per_plex =5,variation_idx = NA,plexes=1)

Histogram_nonOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/Histogram_nonOnePot_MsstatsTMTproc.RDS")
Histogram_non5<-Histogram_nonOnePot_MsstatsTMTproc+theme(text=element_text(size=40))
path<-getwd()
Histogram_non5
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_non5.png")
```
#D. Figure 6, panel F: for a design with 1 TMT 10-plex, 2 biological replicates per mixture and 2 temperatures 

###Simulate a strong interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition
```{r}
OnePot_1plex_2repsstrong<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),design="OnePot",n_replicates_per_plex=2)
Histograms_2temp_strong_onePot_2reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_1plex_2repsstrong,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex=2,variation_idx = NA,plexes=1)

Histogram_strongOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/Histogram_strongOnePot2_MsstatsTMTproc.RDS")
Histogram_strong2<-Histogram_strongOnePot_MsstatsTMTproc+theme(text=element_text(size=40))
path<-getwd()
Histogram_strong2
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_strong2.png")
```
###Simulate a weak interactor for a design with two temperatures, 1 plex, 2 replicates per plex  by Condition

```{r}
OnePot_1plex_2repsweak<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="OnePot",n_replicates_per_plex=2)
Histograms_2temp_weak_onePot_2reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_1plex_2repsweak,design="OnePot",shifter="weak",t_range=seq(5,6),n_replicates_per_plex=2,variation_idx = NA,plexes=1)

Histogram_weakOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_weakOnePot2_MsstatsTMTproc.RDS")
Histogram_weak2<-Histogram_weakOnePot_MsstatsTMTproc+theme(text=element_text(size=40))
path<-getwd()
Histogram_weak2
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_weak2.png")
```

###Simulate a non interactor for a design with two temperatures, 1 plex, 2 replicates per plex  by Condition

```{r}
OnePot_1plex_2repsnon<-benchmark_nonshifter_sim_1plex(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),design="OnePot",n_replicates_per_plex=2)
Histograms_2temp_non_onePot_2reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_1plex_2repsnon,design="OnePot",shifter="non",t_range=seq(5,6),n_replicates_per_plex=2,variation_idx = NA,plexes=1)

Histogram_nonOnePot_MsstatsTMTproc <- readRDS("~/MSstatsThermalProfile/vignettes/Figures5-6_simulation/Histogram_nonOnePot2_MsstatsTMTproc.RDS")
Histogram_non2<-Histogram_nonOnePot_MsstatsTMTproc+theme(text=element_text(size=40))
path<-getwd()
Histogram_non2
ggplot2::ggsave(path = path, width = 10, height = 8, units = 'in', device='png', dpi=600, filename = "Histogram_non2.png")
```
