---
title: "Vignette10"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation_dataset5}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#This vignette outlines a thermal design simulation for figure 5
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggplot2)
library(tidyr)

```

# Simulations

#A. Read in MSstatsTMT summarised protein output:  Dataset 1-Phaneuf *et al.*

Load in MSstatsTMT processed data

```{r}
#read summarised protein output from MSstatsTMT
summarised.proteins<- readRDS("~/MSstatsThermalProfiler/vignettes/Dataset_3a-b_simulation_data/summarised.proteinsD1.RDS")
```

#B.Figure 5, panel D design: 4 TMT 10-plexes, 4 biological replicates and 10 temperatures 
###Print out ICC distributions with MSstatsTMT and simulate a strong shifter for a design with a strong interactor(Panel A)

```{r}
#get variance components
Dataset_1_ICC<-msstats_icc(summarised.proteins,temps=unique(summarised.proteins$ProteinLevelData$temperature))
#generate simulations
Thermal_4plex_4strong<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms for MSstatsTMT
Ten_temps_4plex_4rep<-plot_benchmarks_MSstatsTMT(Thermal_4plex_4strong,design="TPP",shifter="strong")
```
### Simulate a weak shifter(Panel B) 
```{r}

#generate simulations
Thermal_4plex_4rep_weak_10temps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
Ten_temps_4plex_4rep_weak_10temps<-plot_benchmarks_MSstatsTMT(Thermal_4plex_4rep_weak_10temps,design="TPP",shifter="weak")
```

### Print out ICC distributions with MSstatsTMT and simulate a design with a non interactor(Panel C) 
 
```{r}
#read in Zebradata 
summarised.proteinsZebra <- readRDS("~/MSstatsThermalProfiler/vignettes/Dataset_3a-b_simulation_data/summarised.proteinsZebra.RDS")
#generate ICC distribution for all summarised proteins for Dataset 2 with MSstatsTMT processing
Dataset_2_ICC<-msstats_icc(summarised.proteinsZebra,temps=unique(summarised.proteinsZebra$ProteinLevelData$temperature))
#Generate simulations for true negative accession
Thermal_4plex_non<-benchmark_nonshifter_sim(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(1,10),design="TPP")
#plot histograms for MSstatsTMT
Ten_temps_4plex_4rep_non<-plot_benchmarks_MSstatsTMT(Thermal_4plex_non,design="TPP",shifter="non")
```

#C. Figure 5, panel E: for a design with 4 TMT 10-plexes, 1 biological replicate per mixture, 2 temperatures per plex and 2 temperatures as contrast
###Print out ICC distributions and simulate a strong shifter(Panel A) 

```{r}
#generate simulations
Thermal_4plex_1repplex_strong_alltemps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
All_temps_4plex_1repplex_strong<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_1repplex_strong_alltemps,design="TPP",shifter="strong",t_range=seq(1,2),n_replicates_per_plex=1)
```
### Print out ICC distributions and simulate a weak shifter(Panel B) 

```{r}
#generate simulations
Thermal_4plex_1repplex_weak_alltemps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
All_temps_4plex_1repplex_weak<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_1repplex_weak_alltemps,design="TPP",shifter="weak",t_range=seq(1,2),n_replicates_per_plex=1)
```
### Print out ICC distributions and simulate a non shifter(Panel C) 

```{r}
#generate simulations
Thermal_4plex_1repplex_non_alltemps<-benchmark_nonshifter_sim(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
All_temps_4plex_1repplex_non<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_1repplex_non_alltemps,design="TPP",shifter="non",t_range=seq(1,2),n_replicates_per_plex=1)
```

#D. Figure 5, panel F: for a design with 4 TMT 10-plexes, 1 biological replicates per mixture and 2 temperatures as a contrast with variance components from all 10 temperatures 
###Print out ICC distributions and simulate a strong shifter(Panel A)

```{r}
#generate simulations
Thermal_4plex_4rep_strong_2temps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
Two_temps_4plex_4rep_strong<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_4rep_strong_2temps,design="TPP",shifter="strong",t_range=seq(1,10),n_replicates_per_plex=1)
```
### Simulate a weak shifter(Panel B) 

```{r}
#generate simulations
Thermal_4plex_4rep_weak_2temps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(1,10),design="TPP")

#plot histograms
Two_temps_4plex_4rep_weak<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_4rep_weak_2temps,design="TPP",shifter="strong",t_range=seq(1,10),n_replicates_per_plex=1)
```
### Simulate a weak shifter(Panel C) 

```{r}
#generate simulations
Thermal_4plex_4rep_non_2temps<-benchmark_nonshifter_sim(Dataset_2_ICC,
                                                     templateProtein="Q6NV46",
                                                     n_sims=1000,
                                                     t_range=seq(1,10),
                                                     design="TPP")

#plot histograms
Two_temps_4plex_4rep_non<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_4rep_non_2temps,design="TPP",shifter="non",t_range=seq(1,10),n_replicates_per_plex=1)
```

#E.Figure 5, panel G: for a design with two temperatures, 4 plexes, 5 replicates per plex
### Print out ICC distributions and simulate a strong shifter(panel A)

```{r}
#generate simulations
Thermal_4plex_5rep_strong_alltemps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),design="TPP")

#plot histograms
All_temps_4plex_5rep_strong<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_5rep_strong_alltemps,design="TPP",shifter="strong",n_replicates_per_plex=5)
```
### Print out ICC distributions and simulate a weak shifter(panel B)

```{r}
#generate simulations
Thermal_4plex_5rep_weak_2temps<-benchmark_shifter_sim(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="TPP")

#plot histograms
All_temps_4plex_5rep_weak<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_5rep_weak_2temps,design="TPP",shifter="weak",n_replicates_per_plex=5)
```
### Print out ICC distributions and simulate a non shifter(panel C)

```{r}
#generate simulations
Thermal_4plex_5rep_non_2temps<-benchmark_nonshifter_sim(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),design="TPP")

#plot histograms
All_temps_4plex_5rep_non<-plot_benchmarks_MSstatsTMT_reps(Thermal_4plex_5rep_non_2temps,design="TPP",shifter="non",n_replicates_per_plex=5)
```


#F. If comparing with TPP splines, use these functions:
###Compare simulations with TPP statistical model for a design with a strong interactor(Panel A) with 4 TMT 10-plexes, 4 biological replicates and 10 temperatures

```{r}
#plot histograms for TPP
Ten_temps_4plex_4rep_TPP<-plot_benchmarksTPP(Thermal_4plex_4strong,
                                             temps=unique(Thermal_4plex_4strong$temperature),
                                             shifter="strong",
                                             design="TPP")
```
 
###Compare simulations with TPP statistical model for a design with a weak interactor(Panel B) with 4 TMT 10-plexes, 4 biological replicates and 10 temperatures

```{r}
#plot histograms for MSstatsTMT
Ten_weak_4plex_4rep_TPP<-plot_benchmarksTPP(Thermal_4plex_4rep_weak_10temps,
                                             temps=unique(Thermal_4plex_4rep_weak_10temps$temperature),
                                             shifter="weak",
                                             design="TPP")

```

###Compare simulations with TPP statistical model for a design with a non interactor(Panel C) with 4 TMT 10-plexes, 4 biological replicates and 10 temperatures
```{r}

#save histograms for TPP
Ten_temps_4plex_non_TPP<-plot_benchmarksTPP(Thermal_4plex_non,design="TPP",shifter="non")

```
