---
title: "Simulation_dataset3b_OnePot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation_dataset3b_OnePot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#This vignette outlines a thermal pool design simulation for figure 6
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}

library(ggplot2)
library(tidyr)

```

#A. Load in MSstatsTMT processed data

```{r}
#read summarised protein output from MSstatsTMT
summarised.proteins<- readRDS("~/MSstatsThermalProfiler/vignettes/Figures5-6_simulation_data/summarised.proteinsD1.RDS")
summarised.proteinsNS<- readRDS("~/MSstatsThermalProfiler/vignettes/Figures5-6_simulation_data/summarised.proteinsZebra.RDS")
```

#B.Figure 6, panel D design: 2 TMT 10-plexes, 10 biological replicates and 2 temperatures 
###Print out ICC distributions and simulate a strong interactor for a design with two temperatures, 2 plexes, 10 replicates

```{r}
#get variance components
Dataset_1_ICC<-msstats_icc(summarised.proteins,temps=unique(summarised.proteins$ProteinLevelData$temperature))
#generate simulations
 OnePot_2plex_10repsstrong<-benchmark_shifter_sim_2plex(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),design="OnePot",shifter="strong")
Histograms_2temp_strong_onePot_10reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_10repsstrong,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex =5)
```
###Simulate a weak interactor for a design with two temperatures, 2 plexes, 5 replicates per plex  by Condition

```{r}
#generate simulations
 OnePot_2plex_10repsweak<-benchmark_shifter_sim_2plex(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="OnePot",shifter="weak")
Histograms_2temp_weak_onePot_10reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_10repsweak,design="OnePot",shifter="weak",t_range=seq(5,6),n_replicates_per_plex =5)
```
##Print out ICC values and simulate a non-interactor for a design with two temperatures, 2 plexes, 5 replicates per plex by Condition

```{r}
#get variance components
Dataset_2_ICC<-msstats_icc(summarised.proteinsNS,temps=unique(summarised.proteinsNS$ProteinLevelData$temperature))
#generate simulations

 OnePot_2plex_10repsnon<-benchmark_nonshifter_sim_2plex(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_non_onePot_10reps<-plot_benchmarks_MSstatsTMT_reps(  OnePot_2plex_10repsnon,design="OnePot",shifter="non",t_range=seq(5,6),n_replicates_per_plex =5)
```
#C. Figure 6, panel E: for a design with 1 TMT 10-plex, 1 biological replicate per mixture, 2 temperatures per plex and 2 temperatures as contrast
###Simulate a strong interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition
```{r}
OnePot_1plex_5repsstrong<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_strong_onePot_5reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_1plex_5repsstrong,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex =5)
```
###Simulate a weak interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition

```{r}
OnePot_1plex_5repsweak<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_weak_onePot_5reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_5repsweak,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex =5)
```

###Simulate a non interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition

```{r}
OnePot_1plex_5repsnon<-benchmark_nonshifter_sim_1plex(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_non_onePot_5reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_5repsnon,design="OnePot",shifter="non",t_range=seq(5,6),n_replicates_per_plex =5)
```
#D. Figure 6, panel F: for a design with 1 TMT 10-plex, 2 biological replicates per mixture and 2 temperatures 

###Simulate a strong interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition
```{r}
OnePot_1plex_2repsstrong<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="P36507",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_strong_onePot_2reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_1plex_2repsstrong,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex=2)
```
###Simulate a weak interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition

```{r}
OnePot_1plex_2repsweak<-benchmark_shifter_sim_1plex(Dataset_1_ICC,templateProtein="O00422",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_weak_onePot_2reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_2repsweak,design="OnePot",shifter="strong",t_range=seq(5,6),n_replicates_per_plex=2)
```

###Simulate a non interactor for a design with two temperatures, 1 plex, 5 replicates per plex  by Condition

```{r}
OnePot_1plex_2repsnon<-benchmark_nonshifter_sim_1plex(Dataset_2_ICC,templateProtein="Q6NV46",n_sims=1000,t_range=seq(5,6),design="OnePot")
Histograms_2temp_non_onePot_2reps<-plot_benchmarks_MSstatsTMT_reps( OnePot_2plex_2repsnon,design="OnePot",shifter="non",t_range=seq(5,6),n_replicates_per_plex=2)
```
